---
title: "Odds Ratio - Up vs Down"
output:
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bugphyzzAnalyses)
library(bugphyzz)
library(bugsigdbr)
library(dplyr)
library(purrr)
library(tidyr)
library(ComplexHeatmap)
```

## Import BugPhyzz

```{r, warning=FALSE, message=FALSE}
bp <- importBugphyzz()
bp_sigs_gn <- map(bp, ~ {
    makeSignatures(
        dat = .x, tax_id_type = "NCBI_ID", tax_level = "genus", min_size = 5
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null) |> 
    map(as.character)

bp_sigs_sp <- map(bp, ~ {
    makeSignatures(
        dat = .x, tax_id_type = "NCBI_ID", tax_level = "species", min_size = 5
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null) |> 
    map(as.character)
```

## Import data

Use "case-control" study design only:

```{r, message=FALSE}
bsdb_doi <- "10.5281/zenodo.10627578" # v1.2.1
# bsdb_doi <-  "10.5281/zenodo.10407666" #v1.2.0
# bsdb_doi <-  "10.5281/zenodo.6468009" #v1.1.0
bsdb <- importBugSigDB(version = bsdb_doi)
bsdb <- bsdb |> 
    filter(`Host species` == "Homo sapiens") |> 
    filter(!is.na(`Abundance in Group 1`)) |> 
    filter(!is.na(`Body site`)) |> 
    filter(`Study design` == "case-control" ) |> 
    mutate(
        `BSDB ID` = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment), "/",
            sub("Signature ", "", `Signature page name`)
        )
    ) |> 
    mutate(
        custom_id = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment)
        )
    ) |> 
    relocate(custom_id, `BSDB ID`)
dim(bsdb)
```

Create a list with signature ids (BSDB ID) that have both UP and DOWN 
outcomes per study.

Genus signatures:

```{r}
up_gn <- getMetaSignatures(
    df = bsdb, column = "Condition", direction = "UP",
    min.studies = 2, min.taxa = 5, tax.id.type = "ncbi",
    tax.level = "genus", exact.tax.level = FALSE
) |> 
    discard(~ length(.x) < 5) |> 
    map(names)
down_gn <- getMetaSignatures(
    df = bsdb, column =  "Condition", direction = "DOWN",
    min.studies = 2, min.taxa = 5, tax.id.type = "ncbi",
    tax.level = "genus", exact.tax.level = FALSE
) |> 
    discard(~ length(.x) < 5) |> 
    map(names)
common_names_gn <- intersect(names(up_gn), names(down_gn))
up_gn <- up_gn[common_names_gn]
down_gn <- down_gn[common_names_gn]
remove_gn <- map2(up_gn, down_gn, ~ intersect(.x, .y))
up_gn <- map2(up_gn, remove_gn, ~ .x[!.x %in% .y])
down_gn <- map2(down_gn, remove_gn, ~ .x[!.x %in% .y])

up_gn <- discard(up_gn, ~ length(.x) < 5)
down_gn <- discard(down_gn,~ length(.x) < 5)

keep_names <- intersect(names(up_gn), names(down_gn))
up_gn <- up_gn[keep_names]
down_gn <- down_gn[keep_names]
# names(up_gn) <- paste0("genus_UP_", names(up_gn))
# names(down_gn) <- paste0("genus_DOWN_", names(down_gn))
```

Species signatures:

```{r}
up_sp <- getMetaSignatures(
    df = bsdb, column = "Condition", direction = "UP",
    min.studies = 2, min.taxa = 5, tax.id.type = "ncbi",
    tax.level = "species", exact.tax.level = FALSE
) |> 
    discard(~ length(.x) < 5) |> 
    map(names)
down_sp <- getMetaSignatures(
    df = bsdb, column =  "Condition", direction = "DOWN",
    min.studies = 2, min.taxa = 5, tax.id.type = "ncbi",
    tax.level = "species", exact.tax.level = FALSE
) |> 
    discard(~ length(.x) < 5) |> 
    map(names)
common_names_sp <- intersect(names(up_sp), names(down_sp))
up_sp <- up_sp[common_names_sp]
down_sp <- down_sp[common_names_sp]
remove_sp <- map2(up_sp, down_sp, ~ intersect(.x, .y))
up_sp <- map2(up_sp, remove_sp, ~ .x[!.x %in% .y])
down_sp <- map2(down_sp, remove_sp, ~ .x[!.x %in% .y])

up_sp <- discard(up_sp, ~ length(.x) < 5)
down_sp <- discard(down_sp,~ length(.x) < 5)

keep_names <- intersect(names(up_sp), names(down_sp))
up_sp <- up_sp[keep_names]
down_sp <- down_sp[keep_names]
# names(up_sp) <- paste0("species_UP_", names(up_sp))
# names(down_sp) <- paste0("species_DOWN_", names(down_sp))
```

```{r}
## Columns that make BSDB ID: Study, Experiment, Signature page name
# l <- bsdb |> 
#     select(
#         Study, Experiment, `Signature page name`, `Abundance in Group 1`
#     ) |> 
#     mutate(
#         xcol = paste0(`Abundance in Group 1`, " (", `Signature page name`, ")")
#     ) |> 
#     group_by(Study, Experiment) |> 
#     mutate(ycol = sort(paste0(xcol, collapse = "; "))) |> 
#     mutate(n_sig = n()) |> 
#     ungroup() |> 
#     select(-xcol) |> 
#     filter(n_sig == 2) |> 
#     filter(grepl("decreased", ycol), grepl("increased", ycol)) |> 
#     filter(grepl("Signature 1", ycol), grepl("Signature 2", ycol)) |> 
#     mutate(
#         `BSDB ID` = paste0(
#             "bsdb:",
#             sub("Study ", "", Study), "/",
#             sub("Experiment ", "", Experiment), "/",
#             sub("Signature ", "", `Signature page name`)
#         )
#     ) |> 
#     mutate(
#         custom_id = paste0(
#             "bsdb:",
#             sub("Study ", "", Study), "/",
#             sub("Experiment ", "", Experiment)
#         )
#     ) |> 
#     relocate(custom_id, `BSDB ID`) |> 
#     {\(y) split(y, y$`Abundance in Group 1`)}() |> 
#     map(~ pull(arrange(.x, custom_id), `BSDB ID`))
# map(l, length)
```

## Genus

Create signatures for both decrease and increase at the **genus** level. 
Only keep the signature names (stored in a data.frame) that have enough taxa in
both increased and decreased sets.

```{r}
# lg <- map(l, ~ {
#     bsdb |> 
#         filter(`BSDB ID` %in% .x) |> 
#         getSignatures(
#             tax.id.type = "ncbi", tax.level = "genus", exact.tax.level = FALSE,
#             min.size = 5
#         ) |> 
#         keep(~ length(.x) >= 5)
# })
# df_gn_1 <- data.frame(
#     x = paste0(
#         sub("/\\d_.*", "", names(lg$decreased)), "-%-%-", names(lg$decreased)
#     )
# ) |> 
#     separate(col = "x", into = c("custom_id", "sig_down"), sep = "-%-%-")
# 
# df_gn_2 <- data.frame(
#     x = paste0(
#         sub("/\\d_.*", "", names(lg$increased)), "-%-%-", names(lg$increased)
#     )
# ) |> 
#     separate(col = "x", into = c("custom_id", "sig_up"), sep = "-%-%-")
# 
# df_gn <- full_join(df_gn_1, df_gn_2, by = "custom_id") |> 
#     drop_na() |> 
#     mutate(
#         check = sub("bsdb:\\d+/\\d+/\\d+_(.*)_UP", "\\1", sig_up) ==
#            sub("bsdb:\\d+/\\d+/\\d+_(.*)_DOWN", "\\1", sig_down)
#     )
# dim(df_gn)    
```

## Species

Create signatures for both decrease and increase at the **species** level.
Only keep the signature names (stored in a data.frame) that have enough taxa in
both increased and decreased sets.

```{r}
# ls <- map(l, ~ {
#     bsdb |> 
#         filter(`BSDB ID` %in% .x) |> 
#         getSignatures(
#             tax.id.type = "ncbi", tax.level = "species",
#             exact.tax.level = FALSE, min.size = 5
#         ) |> 
#         keep(~ length(.x) >= 5)
# })
# df_sp_1 <- data.frame(
#     x = paste0(
#         sub("/\\d_.*", "", names(ls$decreased)), "-%-%-", names(ls$decreased)
#     )
# ) |> 
#     separate(col = "x", into = c("custom_id", "sig_down"), sep = "-%-%-")
# 
# df_sp_2 <- data.frame(
#     x = paste0(
#         sub("/\\d_.*", "", names(ls$increased)), "-%-%-", names(ls$increased)
#     )
# ) |> 
#     separate(col = "x", into = c("custom_id", "sig_up"), sep = "-%-%-")
# 
# df_sp <- full_join(df_sp_1, df_sp_2, by = "custom_id") |> 
#     drop_na() |> 
#     mutate(
#         check = sub("bsdb:\\d+/\\d+/\\d+_(.*)_UP", "\\1", sig_up) ==
#             sub("bsdb:\\d+/\\d+/\\d+_(.*)_DOWN", "\\1", sig_down)
#     )
# dim(df_sp)
```

## Get lists of BSDB signatures

```{r}
sigs_list_gn <- list()
sigs_list_gn[["up"]] <- up_gn
sigs_list_gn[["down"]] <- down_gn 
# sigs_list_gn[["up"]] <- lg$increased[df_gn$sig_up]
# sigs_list_gn[["down"]] <- lg$decreased[df_gn$sig_down]

sigs_list_sp <- list()
sigs_list_sp[["up"]] <- up_sp
sigs_list_sp[["down"]] <- down_sp 
# sigs_list_sp[["up"]] <- ls$increased[df_sp$sig_up]
# sigs_list_sp[["down"]] <- ls$decreased[df_sp$sig_down]
```

## Odds ratio - function

```{r}
or <- function(x, y, s) {
    a <- sum(x %in% s)
    b <- sum(!x %in% s)
    c <- sum(y %in% s)
    d <- sum(!y %in% s)
    ct <- matrix(
        data = c(a, b, c, d), byrow = TRUE, nrow = 2, ncol = 2,
        dimnames = list(exposure = c("yes", "no"), event = c("yes", "no"))
    )
    # suppressWarnings({
    #     res <- epitools::oddsratio.wald(ct)
    # })
    p_value = fisher.test(x = ct, alternative = "two.sided")$p.value
    
    if (any(ct == 0)) {
        HA <- "*"
        suppressWarnings({
            # res <- epitools::oddsratio.small(ct)
            # p_value = fisher.test(x = ct + 0.5, alternative = "two.sided")$p.value
            res <- epitools::oddsratio.wald(ct + 0.5) # I get a lot of caution messages about not using this funcion
        })
    } else {
        HA <- ""
        suppressWarnings({
            # p_value = fisher.test(x = ct, alternative = "two.sided")$p.value
            res <- epitools::oddsratio.wald(ct)
        })
    }
    
    output <- data.frame(
       odds_ratio = round(res$measure[2, "estimate"], 2),
       lower_ci = round(res$measure[2, "lower"], 2),
       upper_ci = round(res$measure[2, "upper"], 2),
       # p_value = res$p.value[2, "fisher.exact"],
       p_value = p_value,
       # method = method,
       HA = HA,
       up_annotated = a,
       up_noAnnotated = b,
       down_annotated = c,
       down_noAnnotated = d
    )
    return(output)
}
```

## Calculate odds ratio and p values

```{r}
fname <- system.file(
    'extdata', 'condition2category.tsv', package = 'bugphyzzAnalyses',
    mustWork = TRUE
)
cond2cat <- read.table(fname, header = TRUE, sep = '\t')
```

```{r}
gn <- vector("list", length(sigs_list_gn$up) * length(bp_sigs_gn))
counter <- 1
for (i in seq_along(sigs_list_gn$up)) {
    for(j in seq_along(bp_sigs_gn)) {
        odds_ratio <- or(
            x = sigs_list_gn$up[[i]],
            y = sigs_list_gn$down[[i]],
            s = bp_sigs_gn[[j]]
        )
        names(gn)[counter] <- paste0(
            names(sigs_list_gn$up)[i], "-%-%-", names(bp_sigs_gn)[j], "-%-%-genus"
        )
        gn[[counter]] <- odds_ratio
        counter <- counter + 1
    }
}
gn_df <- gn |> 
    bind_rows(.id = "name") |> 
    separate(
        col = "name", into = c("condition", "bp_sig", "rank"), sep = "-%-%-"
    ) |> 
    mutate(
        bp_sig = sub("^bugphyzz:(.*)$", "\\1", bp_sig)
    ) |> 
    arrange(-odds_ratio) |> 
    mutate(condition = tolower(condition)) |> 
    left_join(cond2cat, by = c("condition" = "Condition")) |> 
    filter(!is.na(Category))
dim(gn_df)
```

```{r}
sp <- vector("list", length(sigs_list_sp$up) * length(bp_sigs_sp))
counter <- 1
for (i in seq_along(sigs_list_sp$up)) {
    for(j in seq_along(bp_sigs_sp)) {
        odds_ratio <- or(
            x = sigs_list_sp$up[[i]],
            y = sigs_list_sp$down[[i]],
            s = bp_sigs_sp[[j]]
        )
        names(sp)[counter] <- paste0(
            names(sigs_list_sp$up)[i], "-%-%-", names(bp_sigs_sp)[j], "-%-%-genus"
        )
        sp[[counter]] <- odds_ratio
        counter <- counter + 1
    }
}
sp_df <- sp |> 
    bind_rows(.id = "name") |> 
    separate(
        col = "name", into = c("condition", "bp_sig", "rank"), sep = "-%-%-"
    ) |> 
    mutate(
        bp_sig = sub("^bugphyzz:(.*)$", "\\1", bp_sig)
    ) |> 
    arrange(-odds_ratio) |> 
    mutate(condition = tolower(condition)) |> 
    left_join(cond2cat, by = c("condition" = "Condition")) |> 
    filter(!is.na(Category))
dim(sp_df)
```

## Merge gn and sp

```{r}
mainDF <- bind_rows(gn_df, sp_df) |> 
    mutate(
        fdr = round(p.adjust(p_value, method = "fdr"), 3),
        p_value = round(p_value, 3),
    ) |> 
    relocate(fdr, .after = p_value)
dim(mainDF)
```

## odds ratio - genus - table up

```{r}
genus_table_up <- mainDF |> 
    filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    arrange(-odds_ratio)
myDataTable(genus_table_up, 30)
```

## Odds ratio - genus - heatmap up

```{r, fig.height=9}
gn_mat_up <- genus_table_up |> 
    # filter(rank == "genus") |> 
    # filter(p_value < 0.05, odds_ratio > 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

color_fun <- function(x) {
    circlize::colorRamp2(
        breaks = c(0, seq(1, max(x))),
        colors = c("white", viridis::viridis(max(x)))
    )
}

gn_hp_up <- Heatmap(
    matrix = gn_mat_up,
    col = color_fun(gn_mat_up), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(gn_mat_up),
        gp = gpar(fontsize = 12)
    ),
)
gn_hp_up
```

## odds ratio - genus - table down

```{r}
genus_table_down <- mainDF |> 
    filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    arrange(-odds_ratio)
myDataTable(genus_table_down, 30)
```

## odds ratio - genus - heatmap down

```{r fig.height=9}
gn_mat_down <- genus_table_down |> 
    # filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()
gn_hp_down <- Heatmap(
    matrix = gn_mat_down,
    col = color_fun(gn_mat_down), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(gn_mat_down),
        gp = gpar(fontsize = 12)
    ),
)
gn_hp_down
```

## odds ratio - species - table up

```{r}
species_table_up <- mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    arrange(-odds_ratio)
myDataTable(species_table_up, 30)
```

## Odds ratio - species - heatmap up

```{r, fig.height=9}
sp_mat_up <- species_table_up |> 
    # filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

sp_hp_up <- Heatmap(
    matrix = sp_mat_up,
    col = color_fun(sp_mat_up), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(sp_mat_up),
        gp = gpar(fontsize = 12)
    ),
)
sp_hp_up
```

## Odds ratio - species - table down

```{r}
species_table_down <- mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    arrange(-odds_ratio)

myDataTable(species_table_down, 30)
```

## Odds ratio - species - heatmap down

```{r, fig.height=9}
sp_mat_down <- species_table_down |> 
    # filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

sp_hp_down <- Heatmap(
    matrix = sp_mat_down,
    col = color_fun(sp_mat_down), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(sp_mat_down),
        gp = gpar(fontsize = 12)
    ),
)
sp_hp_down
```

## Session information

```{r}
sessioninfo::session_info()
```
