---
title: "Analysis of the NYCHANES signatures"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugsigdbr)
library(bugphyzz)
library(dplyr)
library(purrr)
```

# Import data

```{r}
bp <- importBugphyzz(version = 'devel', force_download = TRUE)
bsdb <- importBugSigDB(version = 'devel')
tms <- importTidyTMS()
```

+ Study 727 corresponds to the NYCHANES public study.
+ Signature 1 is decreased abundance in smokers.
+ Signature 2 is increased abundance in smokers.

```{r}
study_727 <- bsdb |> 
    filter(`BSDB ID` %in% c('bsdb:727/1/1', 'bsdb:727/1/2'))
sigs727 <- bugsigdbr::getSignatures(
  df = study_727, tax.id.type = 'ncbi', tax.level = 'mixed'
)
decreased <- sigs727$`bsdb:727/1/1_Tobacco-smoke-exposure-measurement:Current-cigarette-smokers_vs_Never-smokers_DOWN`
increased <- sigs727$`bsdb:727/1/2_Tobacco-smoke-exposure-measurement:Current-cigarette-smokers_vs_Never-smokers_UP`
df <- data.frame(
  NCBI_ID = c(increased, decreased),
  Abundance = c(rep('increased', length(increased)), rep('decreased', length(decreased)))
)
head(df)
```

```{r}
aer <- filter(bp, Attribute_group == 'aerophilicity') |> 
  select(NCBI_ID, Attribute) |> 
  mutate(Attribute = sub('^.*:', '', Attribute)) |> 
  mutate(
    Attribute = case_when(
      Attribute == 'obligately_anaerobic' ~ 'anaerobic',
      Attribute == 'microaerophilic' ~ 'aerobic',
      Attribute == 'obligately_aerobic' ~ 'aerobic',
      TRUE ~ Attribute
    )
  ) |> 
  distinct()
head(aer)
```

Remove taxa marked in both groups

```{r}
tbl <- left_join(df, aer, by = 'NCBI_ID')
dups <- tbl$NCBI_ID[which(duplicated(tbl$NCBI_ID))]
taxizedb::taxid2name(dups)
```

```{r}
tbl[tbl$NCBI_ID %in% dups,]
```

```{r}
tbl <- tbl[!tbl$NCBI_ID %in% dups, ]
any(duplicated(tbl$NCBI_ID))
```

Contingency table

```{r}
cont <- epitools::epitable(tbl$Attribute, tbl$Abundance)
cont
```

Chi-square. Test of independence.

> H0: There is no relationship between smoking and aerophilicity in DA microbes.  
H1: There is a relationship between smoking and aerophilicity in DA microbes.

```{r}
chi_res <- chisq.test(
  x = tbl$Attribute, y = tbl$Abundance, simulate.p.value = TRUE
)
chi_res
```


```{r}
tbl2 <- filter(tbl, Attribute %in% c('aerobic', 'anaerobic'))
cont2 <- epitools::epitable(tbl2$Attribute, tbl2$Abundance)
cont2
```

```{r}
chi_res2 <- chisq.test(
  x = tbl2$Attribute, y = tbl2$Abundance, simulate.p.value = TRUE
)
chi_res2
```

```{r}
mouth <- tms |> 
  filter(body_site == 'oralcavity', rank == 'genus')
```


```{r}
increased_bkg <- unique(c(increased, mouth$taxid))
decreased_bkg <- unique(c(decreased, mouth$taxid))
```

```{r}
aer2 <- bp |> 
  filter(grepl('aerophilicity', Attribute))
bp_sigs <- getBugphyzzSignatures(
  df = aer2, tax.id.type = 'NCBI_ID', tax.level = 'genus'
)
```



```{r}
en2 <- runEnrichment(
  bsdb = bsdb,
  bsdb_sets = list(`bsdb:727/1/1` = decreased, `bsdb:727/1/2` = increased),
  background_sets = list(`bsdb:727/1/1` = decreased_bkg, `bsdb:727/1/2` = increased_bkg),
  bp_sigs = bp_sigs
)
data.table::data.table(en2)

```









# Session information

```{r}
sessioninfo::session_info()
```
