---
title: "Data integration and stats"
output:
  html_document:
    code_folding: hide
    toc: true
---

```{r script options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugphyzz)
library(dplyr)
library(taxPPro)
library(ggplot2)
library(forcats)
library(ggpubr)
library(taxPPro)
library(tidyr)
library(ComplexHeatmap)
library(purrr)
library(forcats)
library(ggbreak)
library(tibble)
library(patchwork)
library(grid)
library(cowplot)
```

## Bugphyzz data

```{r import bp, message=FALSE}
discrete_types <- c(
    "binary", "multistate-intersection", "multistate-union"
)
bp <- importBugphyzz(v = 0.5) |> 
    map(~ mutate(.x, NCBI_ID = as.character(NCBI_ID))) |> 
    map(~ {
        attr_type <- unique(.x$Attribute_type)
        if (attr_type %in% discrete_types) {
            df <- .x |> 
                filter(
                    !(Validation < 0.7 & Evidence == "asr")
                )
        } else if (attr_type == "numeric") {
            df <- .x |> 
                filter(
                    !(Validation < 0.5 & Evidence == "asr")
                )
        }
        df
    })
length(bp)
```
```{r}
myDataTable(data.frame(attributes = names(bp)), page_len = length(bp))
```

## Totals

```{r totals}
## Number of unique taxa
n_taxa <- map(bp, ~ .x$NCBI_ID) |> 
    unlist(use.names = FALSE) |> 
    unique() |> 
    length()

## Number of annotations
n_annotations <- map_int(bp, nrow) |> 
    sum()

## Number of physiologies/attribute groups
n_attr <- length(bp)

## Number of sources
n_sources <- map(bp, ~ {
    unique(.x$Attribute_source) |> 
        {\(y) y[!is.na(y)]}()
}) |> 
    unlist(recursive = TRUE, use.names = FALSE) |> 
    unique() |> 
    length()

## Number of annotations from sources
n_annotations_source <- map_int(bp, ~ {
    .x |> 
        filter(!is.na(Attribute_source)) |> 
        nrow()
}) |> 
    sum()

## Number of annotations IBD
n_annotations_ibd <- map_int(bp, ~ {
    .x |> 
        filter(Evidence == "tax") |> 
        nrow()
}) |> 
    sum()

## Number of annotations ASR
n_annotations_asr <- map_int(bp, ~ {
    .x |> 
        filter(Evidence == "asr") |> 
        nrow()
}) |> 
    sum()

## Number of 
num_vs_dis_n <- map_chr(bp, ~ {
    unique(.x$Attribute_type)
}) |> 
    table() |> 
    as.data.frame() |> 
    set_names(c("name", "n")) |> 
    mutate(name = ifelse(name == "numeric", "numeric", "discrete")) |> 
    count(name, wt = n)

attr_asr_default_names <- map(bp, ~ {
    unique(.x$Evidence)
}) |> 
    keep(~ "asr" %in% .x) |> 
    names()

attr_asr_default <- length(attr_asr_default_names)

attr_asr_all <- bugphyzz:::.validationData() |>
    filter(rank == "all") |> 
    select(physiology, mcc_mean, r2_mean) |> 
    pull(physiology) |> 
    unique() |> 
    length()

count_summary_tbl <- tibble::tribble(
    ~name, ~n,
    "Number of taxa", n_taxa,
    "Number of annotations", n_annotations,
    "Number of attributes", n_attr,
    "Number of discrete attributes", pull(filter(num_vs_dis_n, name == "discrete"),n),
    "Number of sources", n_sources,
    "Number of annoations from sources", n_annotations_source,
    "Number of IBD annotations", n_annotations_ibd,
    "Number of annotations imputed through ASR", n_annotations_asr,
    "Number of attributes with ASR data (default > 0.5 val)", attr_asr_default,
    "Number of attributes with ASR data (all)", attr_asr_all
)

myDataTable(count_summary_tbl)
```

## Totals by rank

```{r, fig.width=10, fig.height=6}
valid_ranks <- c(
    "superkingdom", "phylum", "class", "order", "family",
    "genus", "species", "strain"
)
attr_order <- names(sort(map_int(bp, nrow), decreasing = TRUE))
counts_rank <- bp |> 
    map(~ count(.x, Attribute, Rank)) |> 
    bind_rows() |> 
    mutate(Attribute = factor(Attribute, levels = attr_order)) |> 
    mutate(Rank = factor(Rank, levels = valid_ranks))
p_counts_rank <- counts_rank |> 
    ggplot(aes(Attribute, n)) +
    geom_col(aes(fill = Rank), position = "stack") +
    scale_fill_discrete(name = "Rank") +
    scale_y_continuous(labels = scales::comma) +
    labs(
        x = "Attribute", y = "No. annotations"
    ) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
p_counts_rank
```

## Total number of annotations by type of Evidence

```{r}
count_types <- bp |> 
    map (~{
        .x |> 
            mutate(
                Type = case_when(
                    Evidence %in% c("exp", "igc", "nas", "tas") ~ "Source",
                    Evidence == "asr" ~ "ASR",
                    Evidence == "tax" ~ "IBD"
                )
            ) |> 
            count(Attribute, Type)
}) |> 
    bind_rows()


keep_attrs <- split(count_types, count_types$Attribute) |> 
    keep( ~ {
        types <- unique(.x$Type)
        "IBD" %in% types | "ASR" %in% types
    }) |> 
    names()

p_count_types <- count_types |> 
    filter(Attribute %in% keep_attrs) |> 
    mutate(Attribute = factor(Attribute, levels = attr_order)) |> 
    ggplot(aes(Attribute, n)) +
    geom_col(aes(fill = Type), position = "stack") +
    labs(
        x = "Attribute", y = "No. annotations"
    ) +
    scale_fill_discrete(name = "Annotation types") +
    scale_y_continuous(labels = scales::comma) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
p_count_types
```

# Figure: 10-fold cross-validation and counts per evidence type (species)

Create plot of MMC values for discrete attributes:

```{r}
# val_fname <- "/home/samuelgamboa/Projects/taxPProValidation/validation_summary.tsv"
# val <- readr::read_tsv(val_fname, show_col_types = FALSE) |> 
#     filter(rank == "all")

```

```{r validation discrete}
val <- bugphyzz:::.validationData() |>
filter(rank == "all") # Validation including taxa of mixed ranks as input (keyword is "all")
dis_phys_names <- val |> 
    filter(!is.na(mcc_mean)) |> 
    group_by(physiology) |> 
    slice_max(order_by = mcc_mean, n = 1) |> 
    arrange(-mcc_mean) |> 
    pull(physiology)
pValDis <- val |> # plot Validation Discrete
    filter(!is.na(mcc_mean))  |> 
    mutate(ltp_bp = ifelse(is.na(ltp_bp), ltp_bp_phys, ltp_bp)) |> 
    mutate(physiology = factor(physiology, levels = dis_phys_names)) |>
    mutate(
        guide_col = ifelse(
            as.character(physiology) == attribute,
            as.character(physiology),
            paste0(as.character(physiology), "|", attribute)
        )
    ) |> 
    arrange(physiology, mcc_mean) |> 
    ggplot(aes(forcats::fct_inorder(guide_col), mcc_mean, group = attribute)) +
    geom_errorbar(
        aes(ymin = mcc_mean - mcc_sd, ymax = mcc_mean + mcc_sd),
        position = position_dodge(0.5),
        width = 0, linewidth = 0.3
        
    ) +
    geom_hline(
        yintercept = 0.7, linetype = "dotdash", linewidth = 0.2,
        color = "red"
    ) +
    geom_point(
        aes(size = ltp_bp),
        shape = 23, position = position_dodge(0.5), fill = "white"
        # size = 2
    ) +
    scale_size_continuous(name = "tips annotated") +
    scale_y_continuous(
        breaks = round(
            seq(
                min(val$mcc_mean - val$mcc_sd, na.rm = TRUE) - 0.1,
                max(val$mcc_mean + val$mcc_sd, na.rm = TRUE) + 0.1,
                0.2),
            1)
    ) +
    labs(
        y = "Matthews Correlation Coefficient",
        x = "Attribute|Attribute value"
    ) +
    theme_bw()  +
    theme(
        panel.grid.minor = element_blank()
        
    ) +
    coord_flip()
```

Create plot of R-squared values for continuous attributes:

```{r validation numeric}
num <- val |>
    filter(!is.na(r2_mean))
num_phys_names <- num |> 
    group_by(physiology) |> 
    slice_max(order_by = r2_mean, n = 1) |> 
    arrange(-r2_mean) |> 
    pull(physiology)
pValNum <- val |>  # plot Validation Numeric attributes
    filter(!is.na(r2_mean))  |> 
    mutate(ltp_bp = ifelse(is.na(ltp_bp), ltp_bp_phys, ltp_bp)) |> 
    ggplot(aes(reorder(physiology, -r2_mean), r2_mean, group = attribute)) +
    geom_errorbar(
        aes(ymin = r2_mean - r2_sd, ymax = r2_mean + r2_sd),
        position = position_dodge(0.5),
        width = 0, linewidth = 0.3
        
    ) +
    geom_hline(
        yintercept = 0.5, linetype = "dotdash", linewidth = 0.2,
        color = "red"
    ) +
    geom_point(
        aes(size = ltp_bp),
        shape = 23, position = position_dodge(0.5), fill = "white"
        # size = 2
    ) +
    scale_size_continuous(name = "tips annotated") +
    scale_y_continuous(
        breaks = round(seq(min(val$r2_mean - val$r2_sd, na.rm = TRUE) - 0.1, max(val$r2_mean + val$r2_sd, na.rm = TRUE) + 0.1, 0.2), 1)
        
    ) +
    labs(
        y = "R-squared",
        x = "Attribute"
    ) +
    theme_bw()  +
    theme(
        panel.grid.minor = element_blank()
        
    ) +
    coord_flip()
```

Crete plot of counts of annotations at the species level. Stack colors/columns
with colors according to Evidence type of the annotations: ASR, IBD, Source.
Only include attributes that were expanded with IBD or ASR.

```{r counts species}
counts_type_sp <- bp |> 
    map(~ {
        .x |> 
            filter(Rank == "species") |> 
            mutate(
                Evidence = case_when(
                    Evidence == "tax" ~ "IBD",
                    Evidence == "asr" ~ "ASR",
                    Evidence %in% c("exp", "igc", "tas", "nas") ~ "Source",
                    TRUE ~ Evidence
                )
            ) |>
            count(Attribute, Evidence) |> 
            mutate(total = sum(n))
    }) |> 
    bind_rows()
p_counts_sp <- counts_type_sp |> 
    filter(Attribute %in% keep_attrs) |> 
    mutate(Evidence = factor(Evidence, levels = c("ASR", "IBD", "Source"))) |> 
    ggplot(aes(reorder(Attribute, total), n)) +
    geom_col(aes(fill = Evidence), position = "stack") +
    labs(
        y = "No. annotations (species)", x = "Attribute"
    ) +
    scale_fill_manual(
        values = c("#56B4E9", "#D55E00", "#0072B2"),
        name = "Annotation type"
    ) +
    scale_y_continuous(labels = scales::comma) +
    theme_bw() +
    coord_flip()
```

Merge the plots from the validation results and the counts of annotations

```{r validation and counts sp plot, fig.height=8, fig.width=14}
myPlot2 <- plot_grid(
    pValNum, p_counts_sp, ncol = 1,
    labels = c("b)", "c)"),
    rel_heights = c(0.2, 0.8), align = "v"
)
myPlot1 <- plot_grid(
    pValDis, myPlot2,
    labels = c("a)"),
    rel_widths = c(0.55, 0.45), axis = "b"

)
myPlot1
```

# Supplementary Figure: Counts of annotations for genus, spescies, and strain and number of annotations per taxon

Number of annotations at the genus, species, and strain level:

```{r}
counts_type_2 <- bp |> 
    map(~ {
        .x |> 
            filter(Rank %in% c("genus", "species", "strain")) |>
            mutate(
                Evidence = case_when(
                    Evidence == "tax" ~ "IBD",
                    Evidence == "asr" ~ "ASR",
                    Evidence %in% c("exp", "igc", "tas", "nas") ~ "Source",
                    TRUE ~ Evidence
                )
            ) |>
            count(Attribute, Rank, Evidence) |> 
            mutate(total = sum(n))
    }) |> 
    bind_rows()
p_counts_2 <- counts_type_2 |>
    filter(Attribute %in% keep_attrs) |> 
    mutate(Evidence = factor(Evidence, levels = c("ASR", "IBD", "Source"))) |> 
    ggplot(aes(reorder(Attribute, total), n)) +
    geom_col(aes(fill = Evidence), position = "stack") +
    labs(
        y = "No. annotations", x = "Attribute"
    ) +
    scale_fill_manual(
        values = c("#56B4E9", "#D55E00", "#0072B2"),
        name = "Annotation type"
    ) +
    scale_y_continuous(labels = scales::comma) +
    theme_bw() +
    facet_wrap(~Rank, nrow = 1, scales = "free_x") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    coord_flip()
```

Frequency of number of annotations per single taxon:

```{r}
n_annot_tax <- bp |> 
   map(~ {
       attr_type <- unique(.x$Attribute_type)
       .x |> 
           filter(Rank %in% c("strain", "species", "genus")) |> 
           select(Attribute, NCBI_ID, Rank)
   }) |> 
    bind_rows() |> 
    count(NCBI_ID, Rank)
p_n_annot_tax <- n_annot_tax |> 
    ggplot() +
    geom_histogram(aes(x = n), binwidth = 1) +
    labs(
        x = "Annotations per\ntaxon",
        y = "Frequency"
    ) +
    facet_wrap(~Rank, ncol = 1) +
    theme_bw()
```

Merge plots above for supplementary figure:

```{r, fig.width=9, fig.height=4, message=FALSE}
suppN <- plot_grid(
    p_counts_2, p_n_annot_tax,
    labels = c("a)", "b)"),
    rel_widths = c(0.8, 0.20),
    nrow = 1
) 
suppN
```

# Figure: Heatmap Sources x Attribute x Rank



```{r}
mySources <- bp |> 
    map(~ {
        .x |> 
            select(Attribute_source, Confidence_in_curation) |> 
            drop_na()
    }) |> 
    bind_rows() |> 
    distinct()

mySourcesVct <- mySources$Confidence_in_curation
names(mySourcesVct) <- mySources$Attribute_source
```



```{r, fig.width=9, fig.height=15}
source_attribute_rank <- bp |> 
    map(~{
        .x |> 
            filter(!is.na(Attribute_source)) |> 
            count(Attribute, Attribute_source, Rank)
    }) |> 
    bind_rows()

mat <- source_attribute_rank |>
    arrange(Attribute, Rank) |> 
    unite(col = "Attribute|Rank", Attribute, Rank, sep = "|") |> 
    pivot_wider(
        names_from = "Attribute_source", values_from = "n"
    ) |> 
    tibble::column_to_rownames(var = "Attribute|Rank") |> 
    as.data.frame() |> 
    as.matrix()
mat[is.na(mat)] <- 0
mat[mat > 0] <- 1

color_fun <- circlize::colorRamp2(
  breaks = c(0, 1), colors = c('white', "gray20")
)
rankAnnotationsDF <- data.frame(
    Rank = sub("^(.*\\|)(\\w+)$", "\\2", rownames(mat))
)
rank_colors =list(
    Rank = c("genus" = "orange","species" = "#009E73", "strain" = "#CC79A7")
)
row_ha <- rowAnnotation(
    df = rankAnnotationsDF, col = rank_colors, width = unit(5, "cm")
)

mySourcesVct <- mySourcesVct[colnames(mat)]
# mySourcesCol <- case_when(
#     mySourcesVct == "high" ~ "red",
#     mySourcesVct == "medium" ~ "orange",
#     mySourcesVct == "low" ~ "yellow"
# )

mySourcesL <- list(
    Confidence = c(
        high = "red", medium = "orange", low = "yellow"
    )
)


column_ha <- columnAnnotation(
    df = data.frame(
        Confidence = mySourcesVct
    ),
    col = mySourcesL,
    height = unit(5, "cm"),
    annotation_legend_param = list(
        at = c("high", "medium", "low")
    )
)

splits_attrs <- sub("^(.*)\\|\\w+$", "\\1", rownames(mat))
hp <- Heatmap(
    matrix = mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    show_heatmap_legend = TRUE,
    col = color_fun(c(0, 1)),
    name = "Annotation presence",
    left_annotation = row_ha,
    bottom_annotation = column_ha,
    row_split = splits_attrs,
    row_title_rot = 0,
    heatmap_legend_param = list(
        color_bar = "discrete", 
        border = "black"
    ),
    border = TRUE,
    rect_gp = gpar(col = "gray50", lwd = 1),
    column_names_max_height = unit(7, "cm")
)
hp
```

# Table/Sup Table: Sources

```{r}
sources_fname <- system.file(
    "extdata", "attribute_sources.tsv", package = "bugphyzz", mustWork = TRUE
)
sources <- read.table(
    file = sources_fname,
    header = TRUE, sep = "\t",  quote = ""
) |> 
    rename(
        `Source (short)` = Attribute_source,
        `Confidence in curation` = Confidence_in_curation,
        `Source (long)` = full_source
    )
myDataTable(sources, page_len = nrow(sources))
```

## Session information

```{r class.soruce = 'fold-show'}
sessioninfo::session_info()
```
