---
title: "Data integration and bugphyzz stats"
output:
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugphyzz)
library(dplyr)
library(taxPPro)
library(ggplot2)
library(forcats)
library(ggpubr)
library(taxPPro)
library(tidyr)
library(ComplexHeatmap)
library(purrr)
```

Useful variables:

```{r}
valid_ranks <- c(
  'strain', 'species', 'genus', 'family', 'order', 'class', 'phylum',
  'domain'
)
```

# Data

```{r, message=FALSE}
bp <- importBugphyzz(version = 'devel', force_download = TRUE)
```

# Source data

## Number of original sources

```{r}
bp_source <- bp[!is.na(bp$Attribute_source),]
unique_sources <- bp_source |> 
    filter(Evidence  %in% c('exp', 'igc', 'tas', 'nas')) |> 
    pull(Attribute_source) |> 
    unique() |> 
    strsplit(';') |> 
    unlist() |> 
    unique() |> 
    sort()
n_sources <- length(unique_sources)
n_sources
```

## Number of annotaions per source

Here, inheritance and asr are treated as two separate and new sources.
Note that the majority of annotations actually come from the propagation step.

```{r}
summary_source <- bp |> 
  mutate(
        Attribute_source = ifelse(is.na(Attribute_source), Evidence, Attribute_source)
  ) |> 
  mutate(
    Attribute_source = case_when(
      Attribute_source == 'inh' ~ 'Propagation (inheritance)',
      Attribute_source == 'asr' ~ 'Propagation (ASR)',
      TRUE ~ Attribute_source
    )
  ) |> 
  filter(!grepl(';', Attribute_source)) |> 
  count(Attribute_source) |> 
  arrange(-n) |> 
  head(10)
  
# summary_source
summary_source_p <- summary_source |> 
  ggplot(aes(reorder(Attribute_source, -n), n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  labs(
    x = 'Top 10 source', y = '# Annotations'
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2)
  )
summary_source_p
```

Note: These are total numbers. Would it be interesting to see these
by physiology type with percentage?

# Completeness

Import data of taxa used to create the taxonomic tree:

```{r}
base_taxa <- get_ncbi_taxids('bti') |> 
  rename(domain =  kingdom) |> 
  mutate(Rank = ifelse(Rank == 'superkingdom', 'domain', Rank)) |> 
  filter(Rank %in% valid_ranks) |> 
  select(-Taxon_name, -Parent_NCBI_ID)
base_taxa_summary <- base_taxa |> 
  count(Rank, name = "n_inbase")
```
NCBI IDs in data structure

```{r}
## 90% of the taxa in the original taxa can be found in the tree.
bp_original_source <- filter(bp, !is.na(Attribute_source))
mean(bp_original_source$NCBI_ID %in% base_taxa$NCBI_ID)
```


```{r}
## Create sets for calculating completeness
## inset are all ids found in the tree structure
## inset_sources are all ids found in the tree structure and come from 
## original sources only`
inset <- bp |> 
  filter(NCBI_ID %in% base_taxa$NCBI_ID)
inset_summary <- inset |> 
  count(Attribute_group, Rank, name = 'n_inset')
inset_sources <- inset |> 
  filter(!is.na(Attribute_source))
inset_sources_summary <- inset_sources |> 
  count(Attribute_group, Rank, name = 'n_sources')

## Calculate completeness
completeness <- base_taxa_summary |> 
  left_join(inset_summary, by = c('Rank') ) |> 
  left_join(inset_sources_summary, by = c('Attribute_group', 'Rank')) |>
  mutate(
    completeness_before = n_sources / n_inbase,
    completeness_after = n_inset / n_inbase
) |> 
  relocate(
    Attribute_group, Rank, completeness_before, completeness_after,
    n_inbase, n_inset
  )

```

```{r}
completeness <- base_taxa_summary |> 
  left_join(inset_summary, by = c('Rank') ) |> 
  left_join(inset_sources_summary, by = c('Attribute_group', 'Rank')) |>
  mutate(
    completeness_before = n_sources / n_inbase,
    completeness_after = n_inset / n_inbase
) |> 
  relocate(
    Attribute_group, Rank, completeness_before, completeness_after,
    n_inbase, n_inset
  )
```



```{r}
comp_mat_bef <- completeness |> 
  select(Attribute_group, Rank, completeness_before) |> 
  filter(!is.na(completeness_before)) |> 
  pivot_wider(
    names_from = 'Rank', values_from = 'completeness_before',
    values_fill = 0
  ) |> 
  tibble::column_to_rownames(var = 'Attribute_group') |> 
  select(all_of(valid_ranks[1:3])) |> 
  as.matrix() |> 
  round(digits = 2) 

comp_mat_aft <- completeness |> 
  select(Attribute_group, Rank, completeness_after) |> 
  pivot_wider(
    names_from = 'Rank', values_from = 'completeness_after',
    values_fill = 0
  ) |> 
  tibble::column_to_rownames(var = 'Attribute_group') |> 
  select(all_of(valid_ranks)) |> 
  as.matrix() |> 
  round(digits = 2)
comp_mat_bef <- comp_mat_bef[rownames(comp_mat_aft),]
```

Create heatmaps

```{r}
color_fun <- circlize::colorRamp2(
  breaks = c(0, 1), colors = c('white', 'dodgerblue4')
)
lgd <- Legend(col_fun = color_fun, title = "completeness")
ht_b <- Heatmap(
  matrix = comp_mat_bef,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = 'left',
  show_heatmap_legend = FALSE,
  col = color_fun,
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_a <- Heatmap(
  matrix = comp_mat_aft,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = 'left',
  name = 'completeness',
  heatmap_legend_param = list(
    at = c(0, 0.5, 1)
  ),
  col = color_fun,
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_list <- ht_b + ht_a
ht <- draw(
  ht_list, 
  ht_gap = unit(0.4, "cm")
)
```

Add confidence of annotations for this figure.
Discuss the meaning of the inference.
Hold out validation. Choose random annotations (10% ~30-100x). Cross-validation.
Hold out positives and negatives.
One rank or physiology at at a time.
Look for a citation for method implemented for validation.
Check Mathews Correlation coefficient for balanced/unbalanced prediction. Matthews correlation coefficient
Try different tissues.


# Dispersion of sources

Import bugphyzz data before propagation and before resolving conflicts, 
agreements, and double annotations.

```{r import bp spreadsheets}
phys <- physiologies(remove_false = TRUE, full_source = FALSE)
phys_ <- map(phys, ~ prepareDatForPropagation(.x, resolve = FALSE))
```



```{r source summaries (s) by rank and physiologie}
data <- phys_ |> 
  map(~ count(.x, Attribute_source, Rank)) |> 
  bind_rows(.id = 'Attribute_group') |> 
  drop_na() |> 
  filter(Attribute_source != '') |> 
  filter(Attribute_group %in% attr_grps) |> 
  filter(!grepl('BacDive;', Attribute_source)) |> 
  mutate(new_column = paste0(Attribute_group, '|', Rank)) |>
  select(new_column, Attribute_source, n) |> 
  pivot_wider(
    names_from = 'Attribute_source', values_from = 'n', values_fill = 0
  ) |> 
  arrange(new_column) |> 
  tibble::column_to_rownames(var = 'new_column') |> 
  as.matrix()
```


```{r}
# data <- bp_source |> 
#   count(Attribute_source, Attribute_group, Rank) |> 
#   filter(!grepl('BacDive;', Attribute_source)) |> 
#   mutate(new_column = paste0(Attribute_group, '|', Rank)) |>
#   select(new_column, Attribute_source, n) |> 
#   pivot_wider(
#     names_from = 'Attribute_source', values_from = 'n', values_fill = 0
#   ) |> 
#   arrange(new_column) |> 
#   tibble::column_to_rownames(var = 'new_column') |> 
#   as.matrix()
# 

```


```{r}
bar_vct <- sub('^.*\\|', '', rownames(data))
split_vct <- sub('\\|.*$', '', rownames(data))
bar_color <- viridis::viridis(n = 3, option = 'C')
names(bar_color) <- c('genus', 'species', 'strain')
row_ha <- rowAnnotation(
  'Rank' = bar_vct, col = list(Rank = bar_color), 
  show_annotation_name = FALSE
)
log2_data <- log2(data + 1)
ht_sources <- Heatmap(
  matrix = log2_data,
  cluster_rows = FALSE, cluster_columns = FALSE,
  show_row_names = FALSE,
  row_names_side = 'left',
  name = 'log2(# annotations + 1)',
  show_heatmap_legend = TRUE,
  col = color_fun(c(min(log2_data), max(log2_data))),
  left_annotation = row_ha, 
  row_split = split_vct,
  column_names_rot = 45,
  row_title_rot = 0, gap = unit(0.1, 'inches'),
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_sources
```


# Number of taxa

## Top 10 species by number of annotations

This is not really the top ten species since there are a lot of overlaps (ties)
in the number of annotations. So these are the first 10 after arranging
by number of annotations per species and alphabetically.

```{r top 10 species, fig.width=4, fig.height=5}
top10_sp <- bp |> 
  filter(Rank == 'species') |>
  count(NCBI_ID, Taxon_name) |> 
  arrange(-n, Taxon_name) |> 
  head(10)
top10_sp_p <- top10_sp |> 
  ggplot(aes(reorder(Taxon_name, -n), n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  labs(x = 'Top 10 species', y = '# annotations') +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1), 
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2),
    panel.grid.major.x = element_blank()
  )
top10_sp_p
```

## Frequency of number of annotations per taxon

```{r frequency N annotations}
freq_annot_rank <- bp |> 
  count(Rank, NCBI_ID, name = 'n_annot') |> 
  count(Rank, n_annot)
freq_rank_p <- freq_annot_rank |> 
  mutate(
    Rank = factor(Rank, levels = c(
      'strain', 'species', 'genus', 'family', 'order', 'class', 'phylum',
      'domain'
      ))
  ) |> 
  ggplot(aes(n_annot, n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  scale_x_continuous(
    limits = c(0, 11),
    breaks = scales::pretty_breaks()
  ) +
  labs(x = 'Number of annotaions per taxon', y = 'Frequency') +
  facet_wrap(~Rank, scales = 'free_y') +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.1)
  )
freq_rank_p
```

# Numbers per physiology

## Annotations per physiology

```{r, fig.width=4, fig.height=4}
phys_summary <- bp |> 
  count(Attribute_group) |> 
  arrange(-n)
phys_summary_plot <- phys_summary |> 
  ggplot(aes(Attribute_group, n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  labs(y = '# annotations', x = 'Top 10 physiologies') +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2)
  )
phys_summary_plot
```

# Figure 3?

```{r}
fig3 <- ggarrange(
  top10_sp_p,
  phys_summary_plot, labels = c('A)', 'B)'), hjust = 0
)
fig3
```

# Session information

```{r}
sessioninfo::session_info()
```

