---
title: "Data integration and stats"
output:
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```


```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugphyzz)
library(dplyr)
library(taxPPro)
library(ggplot2)
library(forcats)
library(ggpubr)
library(taxPPro)
library(tidyr)
library(ComplexHeatmap)
library(purrr)
```

# Data

```{r, message=FALSE}
bp <- importBugphyzz(version = 'devel')
```

# Source data

## Number of original sources

```{r}
bp_source <- filter(bp, !is.na(Attribute_source))
unique_sources <- bp_source |> 
    filter(Evidence  %in% c('exp', 'igc', 'tas', 'nas')) |> 
    pull(Attribute_source) |> 
    unique() |> 
    strsplit(';') |> 
    unlist() |> 
    unique() |> 
    sort()
n_sources <- length(unique_sources)
n_sources
```

## Number of annotations per source (top 10)

Here, inheritance and asr are treated as two separate and new sources.
Note that the majority of annotations actually come from the propagation step.

```{r source summary}
summary_source <- bp |> 
  mutate(
    Attribute_source = ifelse(
      is.na(Attribute_source), Evidence, Attribute_source
    )
  ) |> 
  mutate(
    Attribute_source = case_when(
      Attribute_source == 'inh' ~ 'Propagation (inheritance)',
      Attribute_source == 'asr' ~ 'Propagation (ASR)',
      TRUE ~ Attribute_source
    )
  ) |> 
  filter(!grepl(';', Attribute_source)) |> # remove repeated annotations for now
  count(Attribute_source) |> 
  arrange(-n) |> 
  head(10) # This is to get only the top 10 of the list
  
# summary_source
summary_source_p <- summary_source |> 
  ggplot(aes(reorder(Attribute_source, -n), n)) +
  geom_col() +
  labs(
    x = 'Top 10 source', y = '# Annotations'
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2)
  )
summary_source_p
```

# Completeness

Import data of taxa used to create the taxonomic tree:

> Completeness sets must be modified. Species in "exclude_all" set + additional species.

```{r}
base_taxa <- get_ncbi_taxids('bti') |> 
  rename(domain =  kingdom) |> 
  mutate(Rank = ifelse(Rank == 'superkingdom', 'domain', Rank)) |> 
  filter(Rank %in% taxRanks()) |> 
  select(-Taxon_name, -Parent_NCBI_ID)
base_taxa_summary <- base_taxa |> 
  count(Rank, name = "n_inbase")
```
NCBI IDs in data structure

```{r}
## 90% of the taxa in the original taxa can be found in the tree.
bp_original_source <- filter(bp, !is.na(Attribute_source))
mean(bp_original_source$NCBI_ID %in% base_taxa$NCBI_ID)
```

```{r}
## Create sets for calculating completeness
## inset are all ids found in the tree structure
## inset_sources are all ids found in the tree structure and come from 
## original sources only`
inset <- bp |> 
  filter(NCBI_ID %in% base_taxa$NCBI_ID) |> 
  select(-Attribute) |> 
  distinct()
inset_summary <- inset |> 
  count(Attribute_group, Rank, name = 'n_inset')
inset_sources <- inset |> 
  filter(!is.na(Attribute_source))
inset_sources_summary <- inset_sources |> 
  count(Attribute_group, Rank, name = 'n_sources')

## Calculate completeness
completeness <- base_taxa_summary |> 
  left_join(inset_summary, by = c('Rank') ) |> 
  left_join(inset_sources_summary, by = c('Attribute_group', 'Rank')) |>
  mutate(
    completeness_before = n_sources / n_inbase,
    completeness_after = n_inset / n_inbase
) |> 
  relocate(
    Attribute_group, Rank, completeness_before, completeness_after,
    n_inbase, n_inset
  )

```

```{r}
# completeness <- base_taxa_summary |> 
#   left_join(inset_summary, by = c('Rank') ) |> 
#   left_join(inset_sources_summary, by = c('Attribute_group', 'Rank')) |>
#   mutate(
#     completeness_before = n_sources / n_inbase,
#     completeness_after = n_inset / n_inbase
# ) |> 
#   relocate(
#     Attribute_group, Rank, completeness_before, completeness_after,
#     n_inbase, n_inset
#   )
```


```{r}
comp_mat_bef <- completeness |> 
  select(Attribute_group, Rank, completeness_before) |> 
  filter(!is.na(completeness_before)) |> 
  pivot_wider(
    names_from = 'Rank', values_from = 'completeness_before',
    values_fill = 0
  ) |> 
  tibble::column_to_rownames(var = 'Attribute_group') |> 
  select(all_of(taxRanks()[1:3])) |> 
  as.matrix() |> 
  round(digits = 2) 

comp_mat_aft <- completeness |> 
  select(Attribute_group, Rank, completeness_after) |> 
  pivot_wider(
    names_from = 'Rank', values_from = 'completeness_after',
    values_fill = 0
  ) |> 
  tibble::column_to_rownames(var = 'Attribute_group') |> 
  select(all_of(taxRanks())) |> 
  as.matrix() |> 
  round(digits = 2)
comp_mat_bef <- comp_mat_bef[rownames(comp_mat_aft),]
```

Create heatmaps

```{r}
color_fun <- circlize::colorRamp2(
  breaks = c(0, 1), colors = c('white', 'dodgerblue4')
)
lgd <- Legend(col_fun = color_fun, title = "completeness")
ht_b <- Heatmap(
  matrix = comp_mat_bef,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = 'left',
  show_heatmap_legend = FALSE,
  col = color_fun,
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_a <- Heatmap(
  matrix = comp_mat_aft,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = 'left',
  name = 'completeness',
  heatmap_legend_param = list(
    at = c(0, 0.5, 1)
  ),
  col = color_fun,
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_list <- ht_b + ht_a
ht <- draw(
  ht_list, 
  ht_gap = unit(0.4, "cm")
)
```

Add confidence of annotations for this figure.
Discuss the meaning of the inference.
Hold out validation. Choose random annotations (10% ~30-100x). Cross-validation.
Hold out positives and negatives.
One rank or physiology at at a time.
Look for a citation for method implemented for validation.
Check Mathews Correlation coefficient for balanced/unbalanced prediction. Matthews correlation coefficient
Try different tissues.


# Dispersion of sources

Import bugphyzz data before propagation and before resolving conflicts, 
agreements, and double annotations.

```{r import bp spreadsheets, message=FALSE}
phys <- physiologies(remove_false = TRUE, full_source = FALSE)
phys_ <- map(phys, ~ prepareDatForPropagation(.x, resolve = FALSE))
```



```{r source summaries (s) by rank and physiologie}
attr_grps <- whichAttrGrp(bp)
data <- phys_ |> 
  map(~ count(.x, Attribute_source, Rank)) |> 
  bind_rows(.id = 'Attribute_group') |> 
  drop_na() |> 
  filter(Attribute_source != '') |> 
  filter(Attribute_group %in% attr_grps) |> 
  filter(!grepl('BacDive;', Attribute_source)) |> 
  mutate(new_column = paste0(Attribute_group, '|', Rank)) |>
  select(new_column, Attribute_source, n) |> 
  pivot_wider(
    names_from = 'Attribute_source', values_from = 'n', values_fill = 0
  ) |> 
  arrange(new_column) |> 
  tibble::column_to_rownames(var = 'new_column') |> 
  as.matrix()
```


```{r, fig.width=6, fig.height=6}
bar_vct <- sub('^.*\\|', '', rownames(data))
split_vct <- sub('\\|.*$', '', rownames(data))
bar_color <- viridis::viridis(n = 3, option = 'C')
names(bar_color) <- c('genus', 'species', 'strain')
row_ha <- rowAnnotation(
  'Rank' = bar_vct, col = list(Rank = bar_color), 
  show_annotation_name = FALSE
)
log2_data <- log2(data + 1)
ht_sources <- Heatmap(
  matrix = log2_data,
  cluster_rows = FALSE, cluster_columns = FALSE,
  show_row_names = FALSE,
  row_names_side = 'left',
  name = 'log2(# annotations + 1)',
  show_heatmap_legend = TRUE,
  col = color_fun(c(min(log2_data), max(log2_data))),
  left_annotation = row_ha, 
  row_split = split_vct,
  column_names_rot = 45,
  row_title_rot = 0, gap = unit(0.1, 'inches'),
  cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "black", fill = NA))
  }
)
ht_sources
```


# Number of taxa

## Top 10 species by number of annotations

This is not really the top ten species since there are a lot of overlaps (ties)
in the number of annotations. So these are the first 10 after arranging
by number of annotations per species and alphabetically.

```{r top 10 species, fig.width=4, fig.height=5}
top10_sp <- bp |> 
  filter(Rank == 'species') |>
  count(NCBI_ID, Taxon_name) |> 
  arrange(-n, Taxon_name) |> 
  head(10)
top10_sp_p <- top10_sp |> 
  ggplot(aes(reorder(Taxon_name, -n), n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  labs(x = 'Top 10 species', y = '# annotations') +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1), 
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2),
    panel.grid.major.x = element_blank()
  )
top10_sp_p
```

## Frequency of number of annotations per taxon

```{r frequency N annotations}
unknowns <- bp |> 
  filter(Evidence %in% c('asr', 'inh') & Frequency == 'unknown')

normal <- bp |> 
  filter(!(Evidence %in% c('asr', 'inh') & Frequency == 'unknown'))

freq_annot_rank1 <- unknowns |> 
  count(Rank, name = 'n_taxa') |> 
  mutate(n_annot = 0)

freq_annot_rank2 <- bp |> 
  count(Rank, NCBI_ID, name = 'n_annot') |> 
  count(Rank, n_annot, name = 'n_taxa')

freq_annot_rank <- bind_rows(freq_annot_rank1, freq_annot_rank2)

freq_rank_p <- freq_annot_rank |> 
  mutate(Rank = factor(Rank, levels = taxRanks())) |> 
  filter(Rank %in% c('genus', 'species', 'strain')) |>
  ggplot(aes(n_annot, n_taxa)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  scale_x_continuous(
    # limits = c(0, 11),
    breaks = scales::pretty_breaks()
  ) +
  labs(x = 'Number of annotaions per taxon', y = 'Frequency') +
  facet_wrap(~Rank, scales = 'free_y') +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.1)
  )
freq_rank_p
```

Let's use log

```{r}
freq_rank_p2 <- freq_annot_rank |> 
  mutate(Rank = factor(Rank, levels = taxRanks())) |> 
  mutate(n_taxa = log2(n_taxa + 1)) |> 
  filter(Rank %in% c('genus', 'species', 'strain')) |>
  ggplot(aes(n_annot, n_taxa)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  scale_x_continuous(
    # limits = c(0, 11),
    breaks = scales::pretty_breaks()
  ) +
  labs(x = 'Number of annotaions per taxon', y = 'Frequency') +
  facet_wrap(~Rank, scales = 'free_y') +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.1)
  )
freq_rank_p2

```

Compare both plots (real numbers and log2)

```{r}
freq_rank_p_combined <- ggarrange(
  freq_rank_p + ggtitle('Real frequencies'), 
  freq_rank_p2 + ggtitle('log2 frequencies'), 
  ncol = 1
)
freq_rank_p_combined 
```




# Numbers per physiology

## Annotations per physiology

```{r, fig.width=4, fig.height=4}
phys_summary <- bp |> 
  count(Attribute_group) |> 
  arrange(-n)
phys_summary_plot <- phys_summary |> 
  ggplot(aes(Attribute_group, n)) +
  geom_col(fill = 'dodgerblue4', alpha = 0.8) +
  labs(y = '# annotations', x = 'Top 10 physiologies') +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = 'black', linewidth = 0.2)
  )
phys_summary_plot
```

# Figure 3?

```{r}
fig3 <- ggarrange(
  top10_sp_p,
  phys_summary_plot, labels = c('A)', 'B)'), hjust = 0
)
fig3
```


# Paragraph

Some calculation for paragraphs


```{r}
bp_subset <- bp |> 
  filter(
    !(Frequency == 'unknown' & Evidence %in% c('asr', 'inh'))
  )
n_annotations <- nrow(bp_subset)
n_unique_taxa <- length(unique(bp_subset$NCBI_ID))
n_unique_attr <- length(unique(bp_subset$Attribute))
n_unique_attr_grp <- length(unique(bp_subset$Attribute_group))
n_annotations_sources <- bp_subset |> 
  filter(!Evidence %in% c('asr', 'inh')) |> 
  nrow()
n_annotations_propagation <- bp_subset |> 
  filter(Evidence %in% c('asr', 'inh')) |> 
  nrow()
n_annotations
```

> bugphyzz contains a total of 
`r format(n_annotations, big.mark = ",", scientific = FALSE)` annotations, 
`r format(n_unique_taxa, big.mark = ",", scientific = FALSE)` unique taxa,
`r format(n_unique_attr, big.mark = ",", scientific = FALSE)` unique attributes, 
and `r format(n_unique_attr_grp, big.mark = ",", scientific = FALSE)` unique attributes groups (traits).
`r format(n_annotations_sources, big.mark = ",", scientific = FALSE)`
(`r round((n_annotations_sources / n_annotations) * 100)` %) annotations come directly
from `r format(n_sources, big.mark = ",", scientific = FALSE)` public sources and the remaining 
`r format(n_annotations_propagation, big.mark = ",", scientific = FALSE)`
(`r round((n_annotations_propagation / n_annotations) * 100)` %) were inferred with our
taxonomic propagation approach (see the ‘taxonomic propagation’ section
in methods), increasing our completeness from an average per taxonomic rank of
*** to *** % (Fig. ***).

# Session information

```{r}
sessioninfo::session_info()
```

