---
title: "Enrichment with dbBact approach"
subtitle: "Filter even signatures then concatenate"
output:
  html_document:
    toc: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugsigdbr)
library(bugphyzz)
library(purrr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(ComplexHeatmap)

## Variables
body_sites <- c(
    skin = "skin", vagina = "vagina", mouth = "mouth", feces = "feces"
)
ranks <- c(genus = "genus", species = "species")
directions <- c(increased = "increased", decreased = "decreased")

score_opt <- "counts"
stat_test <- "wilcox.test"
```

# Data

Import bugphyzz:

```{r, warning=FALSE, message=FALSE}
bp <- importBugphyzz()

## Genus signatures
bpSigs_g <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "genus",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)

## Species signatures
bpSigs_s <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "species",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)
```

Import BugSigDB:

```{r}
bsdb_doi <- "10.5281/zenodo.10627578" # v1.2.1
bsdb <- importBugSigDB(version = bsdb_doi)
bsdb <- bsdb |> 
    filter(`Host species` == "Homo sapiens") |> 
    filter(`Abundance in Group 1` %in% c("increased", "decreased")) |>
    filter(!is.na(`Body site`))
    # filter(`Study design` == "case-control" )
dim(bsdb)
```

```{r}
sort(table(bsdb$`Study design`), decreasing = TRUE)
```

Subset by body site (using ontologies):

```{r}
uberon <- getOntology(onto = "uberon")
bsdb_subsets_by_bodysite <- vector("list", length(body_sites))
names(bsdb_subsets_by_bodysite) <- body_sites
for (i in seq_along(bsdb_subsets_by_bodysite)) {
    if (body_sites[i] == "skin") { ## Didn't find an ontology for skin
        bsdb_subsets_by_bodysite[[i]] <- bsdb |> 
            filter(grepl(body_sites[i], `Body site`, ignore.case = TRUE))
    } else {
        bsdb_subsets_by_bodysite[[i]] <- subsetByOntology(
            bsdb, column = "Body site", term = body_sites[i], ontology = uberon 
        )
    }
}
```

Filter even experiments for genus:

```{r}
even_bsdb_g <- map(
    bsdb_subsets_by_bodysite, ~ filterEvenDat(.x, "genus", min_size = 5)
) |> 
    discard(~ !length(.x))
## Number of experiments
map_int(even_bsdb_g, ~ nrow(.x[[1]]))
```

Types of studies:

```{r}
even_bsdb_g |> 
    list_flatten() |> 
    bind_rows() |> 
    count(`Study design`) |> 
    arrange(-n)
```

Filter even experiments for species:

```{r}
even_bsdb_s <- map(
    bsdb_subsets_by_bodysite, ~ filterEvenDat(.x, "species", min_size = 5)
) |> 
    discard(~ !length(.x))
## Number of experiments
map_int(even_bsdb_s, ~ nrow(.x[[1]]))
```

```{r}
even_bsdb_s |> 
    list_flatten() |> 
    bind_rows() |> 
    count(`Study design`) |> 
    arrange(-n)
```


Filter even experiments for genus:

```{r}
even_sigs_g <- map(
    bsdb_subsets_by_bodysite, ~ filterEvenSignatures(.x, "genus", min_size = 5)
) |> 
    discard(~ !length(.x))
map_lgl(even_sigs_g, ~ {
    all(names(.x$decreased) == names(.x$increased))
})
```

Filter even experiments for species:

```{r}
even_sigs_s <- map(
    bsdb_subsets_by_bodysite, ~ filterEvenSignatures(.x, "species", min_size = 5)
) |> 
    discard(~ !length(.x))
map_lgl(even_sigs_s, ~ {
    all(names(.x$decreased) == names(.x$increased))
})
```

Create concatenated signatures:

```{r}
catSigs_g <- map(even_sigs_g, concatenateEvenSigs)
catSigs_s <- map(even_sigs_s, concatenateEvenSigs)
```

Run enrichment using dbBact approach:

```{r, warning=FALSE}
myFisherTest <- function(control, case) {
    controlInSet <- sum(control[control > 0])
    controlNotInSet <- sum(control == 0)

    caseInSet <- sum(case[case > 0])
    caseNotInSet <- sum(case == 0)
    mat <- matrix(
        data = c(
            caseInSet, caseNotInSet,
            controlInSet, controlNotInSet
        ),
        byrow = TRUE, nrow = 2
    )
    colnames(mat) <- c("annotated", "noAnnotated")
    rownames(mat) <- c("Case", "Control")

    fisher.test(mat, alternative = "two.sided")
}
stat_test <- "myFisherTest"

## There are warnings for the wilcox test
feces_enriched_g <- map2(
    .x = catSigs_g$feces$decreased[5],
    .y = catSigs_g$feces$increased[5],
    ~ dbEn2(
        control = .x, case = .y, term_list = bpSigs_g, opt = score_opt, 
        f = stat_test
    )
)
feces_enriched_s <- map2(
    .x = catSigs_s$feces$decreased,
    .y = catSigs_s$feces$increased,
    ~ dbEn2(
        control = .x, case = .y, term_list = bpSigs_s, opt = score_opt,
        f = stat_test
    )
)
mouth_enriched_g <- map2(
    .x = catSigs_g$mouth$decreased,
    .y = catSigs_g$mouth$increased,
    ~ dbEn2(
        control = .x, case = .y, term_list = bpSigs_g, opt = score_opt, 
        f = stat_test
    )
)
mouth_enriched_s <- map2(
    .x = catSigs_s$mouth$decreased,
    .y = catSigs_s$mouth$increased,
    ~ dbEn2(
        control = .x, case = .y, term_list = bpSigs_s, opt = score_opt,
        f = stat_test
    )
)
```

# Colorectal cancer

```{r, fig.height=15, fig.width=35}
ht <- dbHt(feces_enriched_s$`Colorectal-cancer`)
```
# Smoking behaviour

```{r, fig.width=25, fig.height=15}
ht2 <- dbHt(mouth_enriched_s$`Smoking-behavior`)
```

# COVID-19

```{r, fig.width=30, fig.height=15}
ht3 <- dbHt(feces_enriched_s$`COVID-19`)
```


# A single study: NYCHANEs in BugSigDB

```{r}

```





# Session information

```{r}
sessioninfo::session_info()
```
