---
title: "Enrichment with dbBact approach"
output:
  html_document:
    toc: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugsigdbr)
library(bugphyzz)
library(purrr)
library(dplyr)
library(stringr)
library(tidyr)
library(ComplexHeatmap)
library(ggplot2)

## Variables
body_sites <- c(
    skin = "skin", vagina = "vagina", mouth = "mouth", feces = "feces"
)
ranks <- c(genus = "genus", species = "species")
directions <- c(increased = "increased", decreased = "decreased")

perm_var <- 1000
freq_var <- 1
```

# Data

Import bugphyzz and create signatures at the genus and species levels:

```{r, warning=FALSE, message=FALSE}
discrete_types <- c(
    "binary", "multistate-intersection", "multistate-union"
)
bp <- importBugphyzz(v = 0.5) |> 
    map(~ mutate(.x, NCBI_ID = as.character(NCBI_ID))) |> 
    map(~ {
        attr_type <- unique(.x$Attribute_type)
        if (attr_type %in% discrete_types) {
            df <- .x |> 
                filter(
                    !(Validation < 0.7 & Evidence == "asr")
                )
        } else if (attr_type == "numeric") {
            df <- .x |> 
                filter(
                    !(Validation < 0.5 & Evidence == "asr")
                )
        }
        df
    })
bpSigs_g <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "genus",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)

bpSigs_s <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "species",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)
```

Import BugSigDB:

```{r}
bsdb_doi <- "10.5281/zenodo.10627578" # v1.2.1
bsdb <- importBugSigDB(version = bsdb_doi)
bsdb <- bsdb |> 
    filter(`Host species` == "Homo sapiens") |> 
    filter(`Abundance in Group 1` %in% c("increased", "decreased")) |>
    filter(!is.na(`Body site`)) |> 
    mutate(exp = sub("^(bsdb:\\d+/\\d+)/\\d+", "\\1", `BSDB ID`))
dim(bsdb)
```

Subset by body site:

```{r}
uberon <- getOntology(onto = "uberon")
bsdb_subsets_by_bodysite <- vector("list", length(body_sites))
names(bsdb_subsets_by_bodysite) <- body_sites
for (i in seq_along(bsdb_subsets_by_bodysite)) {
    if (body_sites[i] == "skin") { ## Didn't find an ontology for skin
        bsdb_subsets_by_bodysite[[i]] <- bsdb |> 
            filter(grepl(body_sites[i], `Body site`, ignore.case = TRUE))
    } else {
        bsdb_subsets_by_bodysite[[i]] <- subsetByOntology(
            bsdb, column = "Body site", term = body_sites[i], ontology = uberon 
        )
    }
}
```

```{r}
dats <- map(bsdb_subsets_by_bodysite, filterEvenDat)
map_int(dats, ~ nrow(.x$decreased)) == map_int(dats, ~ nrow(.x$increased))
```

# Table summary 1

```{r}
nexp_summary <- dats |> 
    map(~ {
        .x$increased |> 
            count(Condition, name = "nexp") |> 
            filter(nexp > 5)
    }) |> 
    bind_rows(.id = "bsite") |> 
    group_by(bsite) |> 
    slice_max(order_by = nexp, n = 10) |> 
    ungroup()
myDataTable(nexp_summary)
```


```{r}
row_dat_list <- list()
```



# Obesity - feces

```{r}
obs_dat <- dats$feces$increased |> # It doesn't matter here if it's increased or decreased
    filter(Condition == "Obesity")

## Got these experiments with manual inspection of the data.frame
obs_exps = c(
    "bsdb:101/1",
    "bsdb:149/1",
    "bsdb:33/1",
    "bsdb:338/1",
    "bsdb:338/2",
    "bsdb:345/1",
    "bsdb:347/1",
    "bsdb:348/1",
    "bsdb:348/3",
    "bsdb:350/1",
    "bsdb:350/2",
    "bsdb:351/1",
    "bsdb:354/1",
    "bsdb:355/1",
    "bsdb:36/5",
    "bsdb:362/1",
    "bsdb:367/1",
    "bsdb:369/1",
    "bsdb:371/3",
    "bsdb:373/1",
    "bsdb:375/1",
    "bsdb:44/1",
    "bsdb:48/1",
    "bsdb:48/2",
    "bsdb:68/1",
    "bsdb:68/2",
    "bsdb:68/3",
    "bsdb:68/4",
    "bsdb:68/5",
    "bsdb:68/6",
    "bsdb:70/1",
    "bsdb:70/2",
    "bsdb:70/3",
    "bsdb:73/1",
    "bsdb:73/2",
    "bsdb:732/1",
    "bsdb:74/1",
    "bsdb:74/3",
    "bsdb:847/1",
    "bsdb:96/1",
    "bsdb:99/1"  
    
)
obs_dat_inc <- dats$feces$increased |> 
    filter(exp %in% obs_exps)
obs_dat_dec <- dats$feces$decreased |> 
    filter(exp %in% obs_exps)

length(obs_exps)
```

```{r}
obs_sigs_s <- getPairedSigs(obs_dat_dec, obs_dat_inc, rank = "species", cat = TRUE)
obs_sigs_g <- getPairedSigs(obs_dat_dec, obs_dat_inc, rank = "genus", cat = TRUE)
```

Enrichment at the species level:

```{r, fig.height=12, fig.width=20}
obs_feces_enriched_s <- dbEn2(
    obs_sigs_s$dec, obs_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()

row_dat_list$obs_feces_sp <- obs_feces_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "Obesity",
        Rank = "species"
    )
dbHt(obs_feces_enriched_s, scm = FALSE)
```


```{r, fig.width=20, fig.height=10}
obs_feces_enriched_g <- dbEn2(
    obs_sigs_g$dec, obs_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()

row_dat_list$obs_feces_gn <- obs_feces_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "Obesity",
        Rank = "genus"
    )
dbHt(obs_feces_enriched_g, scm = FALSE)
```

# Parkinson's disease - feces

```{r}
pd <- dats$feces$increased |> 
    filter(Condition == "Parkinson's disease")
# pd |> 
#     relocate(exp, `Group 0 name`, `Group 1 name`) |> 
#     View()
```

```{r}
pd_exps <- c(
    "bsdb:22/1",
    "bsdb:508/1",
    "bsdb:533/1",
    "bsdb:533/2",
    # "bsdb:537/1",
    # "bsdb:537/2",
    # "bsdb:537/3",
    # "bsdb:537/4",
    # "bsdb:537/6",
    "bsdb:537/7",
    "bsdb:537/8",
    "bsdb:554/1",
    "bsdb:560/1",
    "bsdb:568/1",
    # "bsdb:568/3",
    "bsdb:600/1",
    "bsdb:604/1",
    "bsdb:604/2",
    "bsdb:605/1",
    # "bsdb:605/2",
    "bsdb:607/1",
    "bsdb:607/2",
    "bsdb:716/1",
    # "bsdb:716/2",
    # "bsdb:716/3",
    # "bsdb:716/4",
    "bsdb:717/1",
    # "bsdb:717/2",
    "bsdb:717/3",
    # "bsdb:717/4",
    # "bsdb:717/5",
    # "bsdb:717/6",
    "bsdb:719/3",
    "bsdb:722/1",
    "bsdb:723/1",
    "bsdb:723/3",
    "bsdb:723/4",
    "bsdb:723/5",
    "bsdb:731/1",
    "bsdb:731/2",
    "bsdb:731/4",
    # "bsdb:741/4",
    "bsdb:742/1",
    "bsdb:743/1",
    "bsdb:746/1"
)
pd_dat_inc <- dats$feces$increased |> 
    filter(exp %in% pd_exps)
pd_dat_dec <- dats$feces$decreased |> 
    filter(exp %in% pd_exps)
length(pd_exps)
```

Get signatures:

```{r}
pd_sigs_s <- getPairedSigs(
    pd_dat_dec, pd_dat_inc, rank = "species", cat = TRUE
)
pd_sigs_g <- getPairedSigs(
    pd_dat_dec, pd_dat_inc, rank = "genus", cat = TRUE
)
```

Enrichment at the species level:

```{r, fig.height=12, fig.width=20}
pd_feces_enriched_s <- dbEn2(
    pd_sigs_s$dec, pd_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()

row_dat_list$pd_feces_sp <- pd_feces_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "pd",
        Rank = "species"
    )
dbHt(pd_feces_enriched_s, scm = FALSE, col_pad = 20)
```

Enrichment at the genus level:

```{r, fig.height=8, fig.width=12, warning=FALSE}
pd_feces_enriched_g <- dbEn2(
    pd_sigs_g$dec, pd_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$pd_feces_gn <- pd_feces_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "pd",
        Rank = "genus"
    )
dbHt(pd_feces_enriched_g, scm = FALSE, col_pad = 20)
```

# Colorectal cancer - feces

Get experiments and BSDB rows

```{r}
crc_exps <- crc_dat <- dats$feces$increased |> # It doesn't matter here if it's increased or decreased
    filter(Condition == "Colorectal cancer") |>
    filter(
        grepl("control", `Group 0 name`, ignore.case = TRUE) &
            grepl("(colorectal cancer|crc)", `Group 1 name`, ignore.case = TRUE)
    ) |> 
    pull(exp)
crc_dat_inc <- dats$feces$increased |> 
    filter(exp %in% crc_exps)
crc_dat_dec <- dats$feces$decreased |> 
    filter(exp %in% crc_exps)
length(crc_exps)
```

Get signatures:

```{r}
crc_sigs_s <- getPairedSigs(
    crc_dat_dec, crc_dat_inc, rank = "species", cat = TRUE
)
crc_sigs_g <- getPairedSigs(
    crc_dat_dec, crc_dat_inc, rank = "genus", cat = TRUE
)
```

Enrichment at the species level:

```{r, fig.height=12, fig.width=20}
crc_feces_enriched_s <- dbEn2(
    crc_sigs_s$dec, crc_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()

row_dat_list$crc_feces_sp <- crc_feces_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "crc",
        Rank = "species"
    )
dbHt(crc_feces_enriched_s, scm = FALSE, col_pad = 20)
```

Enrichment at the genus level:

```{r, fig.height=8, fig.width=12, warning=FALSE}
crc_feces_enriched_g <- dbEn2(
    crc_sigs_g$dec, crc_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$crc_feces_gn <- crc_feces_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "crc",
        Rank = "genus"
    )
dbHt(crc_feces_enriched_g, scm = FALSE, col_pad = 20)
```

# Antimicrobial agent - feces 

Interesting, but many different antibiotics are provided.
It would have been more interesting if vancomycin would be among them.

```{r}
aa_dat <- dats$feces$increased |> 
    filter(Condition == "Antimicrobial agent")
# aa_dat |>
#     relocate(exp, `Group 0 name`, `Group 1 name`) |>
#     View()
stringr::str_extract(aa_dat$`Group 1 definition`, "\\b\\w+in\\b")
```

# COVID-19 - feces

```{r}
covid_dat <- dats$feces$increased |> # It doesn't matter here if it's increased or decreased
    filter(Condition == "COVID-19")

## Manual inspection
covid_exps <- c(
    "bsdb:734/1",
    "bsdb:428/1",
    "bsdb:441/1",
    "bsdb:464/3",
    "bsdb:481/1",
    "bsdb:486/1",
    "bsdb:486/2",
    "bsdb:486/3",
    "bsdb:487/3",
    "bsdb:496/1",
    "bsdb:511/4",
    "bsdb:511/5",
    "bsdb:453/4",
    "bsdb:782/3"
)

covid_dat_inc <- dats$feces$increased |> 
    filter(exp %in% covid_exps)
covid_dat_dec <- dats$feces$decreased |> 
    filter(exp %in% covid_exps)
```

Concatenate signatures:

```{r}
covid_sigs_s <- getPairedSigs(covid_dat_dec, covid_dat_inc, rank = "species", cat = TRUE)
covid_sigs_g <- getPairedSigs(covid_dat_dec, covid_dat_inc, rank = "genus", cat = TRUE)
```

Enrichment at the species level:

```{r, fig.height=12, fig.width=20}
covid_feces_enriched_s <- dbEn2(
    covid_sigs_s$dec, covid_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()

row_dat_list$covid_feces_sp <- covid_feces_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "covid",
        Rank = "species"
    )
dbHt(covid_feces_enriched_s, scm = FALSE)
```

```{r, fig.height=12, fig.width=20}
covid_feces_enriched_g <- dbEn2(
    covid_sigs_g$dec, covid_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$covid_feces_gn <- covid_feces_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "covid",
        Rank = "genus"
    )
dbHt(covid_feces_enriched_g, scm = FALSE)
```

# IBS - feces

```{r}
ibs_dat <- dats$feces$increased |> # It doesn't matter here if it's increased or decreased
        filter(Condition == "Irritable bowel syndrome")
ibs_exps <- ibs_dat |> 
    pull(exp)
ibs_dat_inc <- dats$feces$increased |> 
    filter(exp %in% ibs_exps)
ibs_dat_dec <- dats$feces$decreased |> 
    filter(exp %in% ibs_exps)

# ibs_dat |> 
#     relocate(exp, `Group 0 name`, `Group 1 name`) |> 
#     View()
```

```{r}
ibs_sigs_s <- getPairedSigs(
    ibs_dat_dec, ibs_dat_inc, rank = "species", cat = TRUE
)
ibs_sigs_g <- getPairedSigs(
    ibs_dat_dec, ibs_dat_inc, rank = "genus", cat = TRUE
)
```

```{r, fig.height=15, fig.width=20}
ibs_feces_enriched_s <- dbEn2(
    ibs_sigs_s$dec, ibs_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$ibs_feces_sp <- ibs_feces_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "ibs",
        Rank = "species"
    )
dbHt(ibs_feces_enriched_s, scm = FALSE, col_pad = 20)
```

```{r, fig.width=25, fig.height=15}
ibs_feces_enriched_g <- dbEn2(
    ibs_sigs_g$dec, ibs_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$ibs_feces_gn <- ibs_feces_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "feces",
        Condition = "ibs",
        Rank = "genus"
    )
dbHt(ibs_feces_enriched_g, scm = FALSE, col_pad = 20)
```

# Smoking - mouth

```{r}
smo_dat <- dats$mouth$increased |> # It doesn't matter here if it's increased or decreased
        filter(Condition == "Smoking behavior")
smo_exps <- smo_dat |> 
    filter(exp != "bsdb:368/1") |> 
    pull(exp)
smo_dat_inc <- dats$mouth$increased |> 
    filter(exp %in% smo_exps)
smo_dat_dec <- dats$mouth$decreased |> 
    filter(exp %in% smo_exps)
```

Get concatenated signatures:

```{r}
smo_sigs_s <- getPairedSigs(
    smo_dat_dec, smo_dat_inc, rank = "species", cat = TRUE
)
smo_sigs_g <- getPairedSigs(
    smo_dat_dec, smo_dat_inc, rank = "genus", cat = TRUE
)
```

```{r, fig.height=12, fig.width=20}
smo_mouth_enriched_s <- dbEn2(
    smo_sigs_s$dec, smo_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$smo_mouth_sp <- smo_mouth_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "mouth",
        Condition = "smoking",
        Rank = "species"
    )
dbHt(smo_mouth_enriched_s, scm = FALSE, col_pad = 20)
```

```{r, fig.height=12, fig.width=20}
smo_mouth_enriched_g <- dbEn2(
    smo_sigs_g$dec, smo_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$smo_mouth_gn <- smo_mouth_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "mouth",
        Condition = "smoking",
        Rank = "genus"
    )
dbHt(smo_mouth_enriched_g, scm = FALSE, col_pad = 20)
```

# Human papilloma virus infection - vagina

```{r}
hpv_dat <- dats$vagina$increased |> 
    filter(Condition == "Human papilloma virus infection") |> 
    filter(grepl("HPV\\+",  `Group 1 name`))

hpv_exps <- hpv_dat |> 
    pull(exp)

hpv_dat_inc <- dats$vagina$increased |> 
    filter(exp %in% hpv_exps)
hpv_dat_dec <- dats$vagina$decreased |> 
    filter(exp %in% hpv_exps)

# hpv_dat |>
#     relocate(exp, `Group 0 name`, `Group 1 name`) |>
#     View()
```

Get concatenated signatures:

```{r}
hpv_sigs_s <- getPairedSigs(
    hpv_dat_dec, hpv_dat_inc, rank = "species", cat = TRUE
)
hpv_sigs_g <- getPairedSigs(
    hpv_dat_dec, hpv_dat_inc, rank = "genus", cat = TRUE
)
```

Enrichment at the species level:

```{r, fig.height=15, fig.width=20}
hpv_vagina_enriched_s <- dbEn2(
    hpv_sigs_s$dec, hpv_sigs_s$inc, term_list = bpSigs_s,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$hpv_vagina_sp <- hpv_vagina_enriched_s |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "vagina",
        Condition = "hpv",
        Rank = "species"
    )
dbHt(hpv_vagina_enriched_s, scm = TRUE, col_pad = 20)
```

Enrichment at the genus level:

```{r, fig.height=15, fig.width=20}
hpv_vagina_enriched_g <- dbEn2(
    hpv_sigs_g$dec, hpv_sigs_g$inc, term_list = bpSigs_g,
    perm = perm_var, freq = 1
) |> 
    filterDbEnRes()
row_dat_list$hpv_vagina_gn <- hpv_vagina_enriched_g |> 
    rowData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "bp_sig") |> 
    as_tibble() |> 
    mutate(
        Bsite = "vagina",
        Condition = "hpv",
        Rank = "genus"
    )
dbHt(hpv_vagina_enriched_g, scm = TRUE, col_pad = 20)
```


# Summary table

```{r}
sum_tbl <- row_dat_list |> 
    bind_rows() |> 
    select(-PermP) |> 
    mutate(
        P_value = round(P_value, 3),
        FDR = round(FDR, 3),
        Effect_size = round(Effect_size, 2)
    )
    # mutate(
    #     FDR = p.adjust(P_value, method = "fdr"),
    #     PermP_FDR = p.adjust(PermP, method = "fdr")
    # ) |> 
    # filter(P_value < 0.1)
myDataTable(sum_tbl)
```


```{r}
sum_tbl |> 
    ggplot(aes(P_value)) +
    geom_histogram(binwidth = 0.1)
```

```{r}
sum_tbl |> 
    ggplot(aes(P_value)) +
    geom_histogram(binwidth = 0.1) +
    facet_grid(Condition ~ Rank) +
    scale_y_continuous(breaks = seq(0, 8, 2))
```

# Session information

```{r}
sessioninfo::session_info()
```
