---
title: "Enrichment with dbBact scores (concatenated BugSigDB signatures)"
output:
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bugsigdbr)
library(bugphyzz)
library(purrr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

body_sites <- c(
    skin = "skin", vagina = "vagina", mouth = "mouth", feces = "feces"
)
ranks <- c(genus = "genus", species = "species")
directions <- c(increased = "increased", decreased = "decreased")
```

# Data

Import bugphyzz:

```{r, warning=FALSE, message=FALSE}
bp <- importBugphyzz()
bpSigs_g <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "genus",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)

## Add score of 1
bpSigs_g <- map(bpSigs_g, ~ {
    scores <- rep(1, length(.x))
    names(scores) <- .x
    attr(.x, "Scores") <- scores
    .x
})

bpSigs_s <- map(bp, ~ {
    makeSignatures(
        dat = .x,  tax_id_type = "NCBI_ID", tax_level = "species",
        min_size = 10
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null)

## Add score of 1
bpSigs_s <- map(bpSigs_s, ~ {
    scores <- rep(1, length(.x))
    names(scores) <- .x
    attr(.x, "Scores") <- scores
    .x
})
```

A function for calculating dbScores based on the dbBact paper:

```{r}
dbEn <- function(s1, s2, t) {
    size1 <- length(s1)
    size2 <- length(s2)
    scores1 <- attr(t, "Scores")[s1] |>
        {\(y) y[!is.na(y)]}()
    scores2 <- attr(t, "Scores")[s2] |>
        {\(y) y[!is.na(y)]}()
    (sum(scores1) / size1) - (sum(scores2) / size2)
}
```

Import BugSigDB

```{r}
bsdb_doi <- "10.5281/zenodo.10627578" # v1.2.1
# bsdb_doi <-  "10.5281/zenodo.10407666" #v1.2.0
# bsdb_doi <-  "10.5281/zenodo.6468009" #v1.1.0
bsdb <- importBugSigDB(version = bsdb_doi)
bsdb <- bsdb |> 
    filter(`Host species` == "Homo sapiens") |> 
    filter(!is.na(`Abundance in Group 1`)) |> 
    filter(`Abundance in Group 1` %in% c("increased", "decreased")) |>
    filter(!is.na(`Body site`)) |> 
    filter(`Study design` == "case-control" )
```

Subset by body site

```{r}
uberon <- getOntology(onto = "uberon")
bsdb_subsets_by_bodysite <- vector("list", length(body_sites))
names(bsdb_subsets_by_bodysite) <- body_sites
for (i in seq_along(bsdb_subsets_by_bodysite)) {
    if (body_sites[i] == "skin") {
        bsdb_subsets_by_bodysite[[i]] <- bsdb |> 
            filter(grepl(body_sites[i], `Body site`, ignore.case = TRUE))
    } else {
        bsdb_subsets_by_bodysite[[i]] <- subsetByOntology(
            bsdb, column = "Body site", term = body_sites[i], ontology = uberon 
        )
    }
}
```

# Genus

```{r}
sigs_by_subset_g <- map(bsdb_subsets_by_bodysite, ~ {
    bsdb_ids <- getSignatures(
        df = .x, tax.id.type = "ncbi", tax.level = "genus", min.size = 10
    )
    if (!length(bsdb_ids)) {
        return(NULL)
    }
    bsdb_ids <- bsdb_ids |> 
        names() |>
        sub("^(bsdb:\\d+/\\d+/\\d+)_.*$", "\\1", x = _)
    dats <- bsdb |>
        filter(`BSDB ID` %in% bsdb_ids) |>
        group_by(Study, Experiment) |>
        mutate(count = n()) |>
        ungroup() |>
        filter(count == 2) |>
        group_by(Study, Experiment) |>
        arrange(`Abundance in Group 1`) |>
        mutate(comb = paste0(sort(`Abundance in Group 1`), collapse = "-")) |>
        ungroup() |>
        filter(comb == "decreased-increased") |>
        arrange(`BSDB ID`) |>
        {\(y) split(y, y$`Abundance in Group 1`)}()
    if (!length(dats)) {
        return(NULL)
    }
    sigs <- dats |>
        map(~ {
            getSignatures(
                df = .x, tax.id.type = "ncbi", tax.level = "genus", min.size = 10
            )
        })
    names(sigs$decreased) <- sub(
        # "^(bsdb:\\d+/\\d+)/\\d+_.*$", "\\1", names(sigs$decreased)
        "^(bsdb:\\d+/\\d+)/\\d+_(.+):.*$", "\\1_\\2", names(sigs$decreased)
    )
    names(sigs$increased) <- sub(
        # "^(bsdb:\\d+/\\d+)/\\d+_.*$", "\\1", names(sigs$increased)
        "^(bsdb:\\d+/\\d+)/\\d+_(.+):.*$", "\\1_\\2", names(sigs$increased)
    )
    sigs
}) |> 
    discard(is.null)

## Check that all names match
map_lgl(sigs_by_subset_g, ~ {
    all(names(.x$decreased) == names(.x$increased))
})

```

Concatenate BugSigDB signatures by condition

```{r}
sigs_by_subset_cat_g <- map(sigs_by_subset_g, ~ {
    cond_names <- unique(sub("^bsdb:\\d+/\\d_", "", names(.x$decreased)))
    decreased <- vector("list", length(cond_names))
    increased <- vector("list", length(cond_names))
    for (i in seq_along(cond_names)) {
        rgx <- paste0("/\\d+_", cond_names[i], "$")
        select_var <- grep(rgx, names(.x$decreased), value = TRUE)

        names(decreased)[i] <- cond_names[i]
        decreased[[i]] <- unlist(.x$decreased[select_var], use.names = FALSE)

        names(increased)[i] <- cond_names[i]
        increased[[i]] <- unlist(.x$increased[select_var], use.names = FALSE)
    }
    list(decreased = decreased, increased = increased)
})
```

Quick comparison of number of signatures before and after concatenation:

```{r}
gsig <- sigs_by_subset_g |> list_flatten() |> map_int(length)
gcatsig <- sigs_by_subset_cat_g |> list_flatten() |> map_int(length)
data.frame(sig = gsig, catsig = gcatsig)
```

Run enrichment

```{r}
gres <- vector("list", length(bpSigs_g))
names(gres) <- names(bpSigs_g)
for (i in seq_along(gres)) {
    gres[[i]] <- map(sigs_by_subset_cat_g, function(bsite) {
        map2(bsite$increased, bsite$decreased, ~ {
                dbEn(s1 = .x, s2 = .y, t = bpSigs_g[[i]])
        })
    })
}
g <- list_flatten(gres, name_spec = "{outer}|---|{inner}") |> 
    list_flatten(name_spec = "{outer}|---|{inner}") |> 
    unlist(use.names = TRUE)
genDF <- data.frame(
    x = names(g),
    score = unname(g)
) |>  
    separate(
        col = x, into = c("bugphyzz", "body_site", "bsdb"), sep = "\\|---\\|"
    ) |> 
    mutate(bugphyzz = sub("bugphyzz:", "", bugphyzz)) |> 
    mutate(direction = case_when(
        score == 0 ~ "Unchanged",
        score > 0 ~ "Increased", 
        score < 0 ~ "Decreased"
    ))
```

```{r}
genDF |> 
    ggplot(aes(score)) +
    geom_histogram(binwidth = 0.1) +
    labs(y = "Frequency", x = "Enrichment scores") +
    theme_bw()
```

```{r, fig.width=9, fig.height=12}
gmin <- genDF |> 
    filter(direction == "Decreased") |> 
    group_by(body_site) |> 
    slice_max(order_by = score, n = 20) |> 
    ungroup() |> 
    arrange(body_site, score)
gmax <- genDF |> 
    filter(direction == "Increased") |> 
    group_by(body_site) |> 
    slice_min(order_by = score, n = 20) |>
    ungroup() |> 
    arrange(body_site, -score)
gtop <- bind_rows(gmin, gmax)
gtop |> 
    arrange(body_site, score) |> 
    mutate(label = paste0(bugphyzz, "|||", bsdb)) |> 
    mutate(
        direction = factor(direction, levels = c("Increased", "Decreased"))
    ) |> 
    mutate(label = forcats::fct_inorder(label)) |> 
    ggplot(aes(label, score)) +
    geom_col(aes(fill = direction)) +
    labs(
        x = "bugphyzz||bugsigdb", y = "dbBact Score",
        title = "Top 20 dbBact enrichment scores - genus"
    ) +
    facet_wrap(~ body_site, nrow = 2, scales = "free") +
    theme_bw() + 
    coord_flip()
```

# Species

```{r}
sigs_by_subset_s <- map(bsdb_subsets_by_bodysite, ~ {
    bsdb_ids <- getSignatures(
        df = .x, tax.id.type = "ncbi", tax.level = "species", min.size = 10
    )
    if (!length(bsdb_ids)) {
        return(NULL)
    }
    bsdb_ids <- bsdb_ids |> 
        names() |>
        sub("^(bsdb:\\d+/\\d+/\\d+)_.*$", "\\1", x = _)
    dats <- bsdb |>
        filter(`BSDB ID` %in% bsdb_ids) |>
        group_by(Study, Experiment) |>
        mutate(count = n()) |>
        ungroup() |>
        filter(count == 2) |>
        group_by(Study, Experiment) |>
        arrange(`Abundance in Group 1`) |>
        mutate(comb = paste0(sort(`Abundance in Group 1`), collapse = "-")) |>
        ungroup() |>
        filter(comb == "decreased-increased") |>
        arrange(`BSDB ID`) |>
        {\(y) split(y, y$`Abundance in Group 1`)}()
    if (!length(dats)) {
        return(NULL)
    }
    sigs <- dats |>
        map(~ {
            getSignatures(
                df = .x, tax.id.type = "ncbi",
                tax.level = "species", min.size = 10
            )
        })
    names(sigs$decreased) <- sub(
        # "^(bsdb:\\d+/\\d+)/\\d+_.*$", "\\1", names(sigs$decreased)
        "^(bsdb:\\d+/\\d+)/\\d+_(.+):.*$", "\\1_\\2", names(sigs$decreased)
    )
    names(sigs$increased) <- sub(
        # "^(bsdb:\\d+/\\d+)/\\d+_.*$", "\\1", names(sigs$increased)
        "^(bsdb:\\d+/\\d+)/\\d+_(.+):.*$", "\\1_\\2", names(sigs$increased)
    )
    sigs
}) |> 
    discard(is.null)

## Check that all names match
map_lgl(sigs_by_subset_s, ~ {
    all(names(.x$decreased) == names(.x$increased))
})

```

Concatenate BugSigDB signatures by condition

```{r}
sigs_by_subset_cat_s <- map(sigs_by_subset_s, ~ {
    cond_names <- unique(sub("^bsdb:\\d+/\\d_", "", names(.x$decreased)))
    decreased <- vector("list", length(cond_names))
    increased <- vector("list", length(cond_names))
    for (i in seq_along(cond_names)) {
        rgx <- paste0("/\\d+_", cond_names[i], "$")
        select_var <- grep(rgx, names(.x$decreased), value = TRUE)

        names(decreased)[i] <- cond_names[i]
        decreased[[i]] <- unlist(.x$decreased[select_var], use.names = FALSE)

        names(increased)[i] <- cond_names[i]
        increased[[i]] <- unlist(.x$increased[select_var], use.names = FALSE)
    }
    list(decreased = decreased, increased = increased)
})
```

Quick comparison of number of signatures before and after concatenation:

```{r}
ssig <- sigs_by_subset_s |> list_flatten() |> map_int(length)
scatsig <- sigs_by_subset_cat_s |> list_flatten() |> map_int(length)
data.frame(sig = ssig, catsig = scatsig)
```
Run enrichment

```{r}
sres <- vector("list", length(bpSigs_s))
names(sres) <- names(bpSigs_s)
for (i in seq_along(sres)) {
    sres[[i]] <- map(sigs_by_subset_cat_s, function(bsite) {
        map2(bsite$increased, bsite$decreased, ~ {
                dbEn(s1 = .x, s2 = .y, t = bpSigs_s[[i]])
        })
    })
}
s <- list_flatten(sres, name_spec = "{outer}|---|{inner}") |> 
    list_flatten(name_spec = "{outer}|---|{inner}") |> 
    unlist(use.names = TRUE)
senDF <- data.frame(
    x = names(s),
    score = unname(s)
) |> 
    separate(
        col = x, into = c("bugphyzz", "body_site", "bsdb"), sep = "\\|---\\|"
    ) |> 
    mutate(bugphyzz = sub("bugphyzz:", "", bugphyzz)) |> 
    mutate(direction = case_when(
        score == 0 ~ "Unchanged",
        score > 0 ~ "Increased", 
        score < 0 ~ "Decreased"
    ))
```

```{r}
senDF |> 
    ggplot(aes(score)) +
    geom_histogram(binwidth = 0.1) +
    labs(y = "Frequency", x = "Enrichment scores") +
    theme_bw()
```

```{r, fig.width=9, fig.height=12}
smin <- senDF |> 
    filter(direction == "Decreased") |> 
    group_by(body_site) |> 
    slice_max(order_by = score, n = 20) |> 
    ungroup() |> 
    arrange(body_site, score)
smax <- senDF |> 
    filter(direction == "Increased") |> 
    group_by(body_site) |> 
    slice_min(order_by = score, n = 20) |>
    ungroup() |> 
    arrange(body_site, -score)
stop <- bind_rows(smin, smax)
stop |> 
    arrange(body_site, score) |> 
    mutate(label = paste0(bugphyzz, "|||", bsdb)) |> 
    mutate(
        direction = factor(direction, levels = c("Increased", "Decreased"))
    ) |> 
    mutate(label = forcats::fct_inorder(label)) |> 
    ggplot(aes(label, score)) +
    geom_col(aes(fill = direction)) +
    labs(
        x = "bugphyzz||bugsigdb", y = "dbBact Score",
        title = "Top 20 dbBact enrichment scores - species"
    ) +
    facet_wrap(~ body_site, nrow = 2, scales = "free") +
    theme_bw() + 
    coord_flip()
```

# Session information

```{r}
sessioninfo::session_info()
```
