---
title: "Enrichment of facultative anaerobes in IBD samples"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugphyzz)
library(curatedMetagenomicData)
library(dplyr)
library(tidySummarizedExperiment)
library(taxPPro)
library(EnrichmentBrowser)
library(ggplot2)
```

# Expected result

+ Enrichment of anaerobic bacteria in healthy samples (gut).
+ Enrichment of facultative anaerobic bacteria in IBD samples (gut).

# Data

## Abundance data

For this example, let's use the data of the HMP_2019_bidmdb study through cMD:

```{r, message=FALSE, warning=FALSE}
tse <- curatedMetagenomicData(
    pattern = "HallAB_2017.relative_abundance", 
    dryrun = FALSE, rownames = 'NCBI',
    counts = TRUE
)[[1]]
tse 
```

Use only a subset:

```{r, message=FALSE, warning=FALSE}
rank_cols <- colnames(rowData(tse))
tidy_data <- tidySummarizedExperiment::as_tibble(tse) |> 
    rename(taxon_name = .feature, sample = .sample) 
sample_data <- tidy_data |> 
    select(-rank_cols, -taxon_name, -relative_abundance) |> 
    distinct()
sample_data_split <- split(sample_data, factor(sample_data$disease))
min_n <- min(vapply(sample_data_split, nrow, integer(1)))
samples <- lapply(sample_data_split, function(x) {
    set.seed(1234)
    int_vct <- sample(1:nrow(x), size = min_n)
    x[int_vct,]$sample
}) |> 
    unlist(use.names = FALSE)
tse_subset <- tse[,samples] |> 
    filterTaxa()
tse_subset
```

## Signatures and annotations

```{r}
taxa_data <- tidy_data |> 
    dplyr::filter(taxon_name %in% rownames(tse_subset)) |> 
    select(all_of(c('taxon_name', rank_cols))) |> 
    distinct()
dim(taxa_data)
```

Annotations and microbe signatures come from the bugphyzz package. Since
cMD contains data at the species level, let's use signatures at the species
level. We need to check the completness (number of taxa in cMD samples
annotated in bugphyzz)

```{r, message=FALSE}
aer <- physiologies('aerophilicity')[[1]]
aer_species <- filter(aer, Rank == 'species')
## Check completness
mean(taxa_data$taxon_name %in% aer_species$NCBI_ID) * 100
```

Let's try to increase the completeness with ASR/inheritance propagation using
majority vote:

```{r, message=FALSE, warning=FALSE}
aer_plus <- aer |> 
    propagate()
aer_plus_species <- filter(aer_plus, Rank == 'species')
## Check completeness
mean(taxa_data$taxon_name %in% aer_plus_species$NCBI_ID) * 100
```

Let's get some signatures for aerobic, anaerobic, and facultatively anaerobic:

```{r, message=FALSE, warning=FALSE}
attribute_names <- c('aerobic', 'anaerobic', 'facultatively anaerobic')
aer_sig <- getSignatures(
    aer, tax.id.type = 'NCBI_ID', tax.level = 'species'
)
aer_sig <- aer_sig[names(aer_sig) %in% attribute_names]
aer_sig <- lapply(aer_sig, as.character)
lapply(aer_sig, head)
```

For propagated signatures

```{r, message=FALSE, warning=FALSE}
aer_plus_sig <- getSignatures(
    aer_plus, tax.id.type = 'NCBI_ID', tax.level = 'species'
)
aer_plus_sig <- aer_plus_sig[names(aer_plus_sig) %in% attribute_names]
aer_plus_sig <- lapply(aer_plus_sig, as.character)
lapply(aer_plus_sig, head)
```

Comparing numbers

```{r}
df <- data.frame(
    aer = vapply(aer_sig, length, integer(1)),
    aer_plus = vapply(aer_plus_sig, length, integer(1))
)
df
```

## Add annotations to rowData

```{r}
taxa_data <- as.data.frame(rowData(tse_subset))
```


We got a lot more species

# Differential abundance analysis

Let's use `EnrichmentBrowser` for this

```{r}
## Let's add a column with groups 0 and 1
colData(tse_subset)$GROUP <- 
    ifelse(colData(tse_subset)$disease == 'healthy', 0, 1)

assay(tse_subset) <- assay(tse_subset) + 1 

da_output <- deAna(
    expr = tse_subset, de.method = 'DESeq2', padj.method = 'BH' 
)
dim(da_output)
```


# Enrichment Analysis

```{r, message=FALSE, warning=FALSE}
enrich_output <- sbea(
    method = 'gsea', se = da_output, gs = aer_sig, alpha = 0.05, perm = 0,
    padj.method = 'none'
)
gsRanking(res = enrich_output, signif.only = TRUE)
```

```{r}
enrich_output_plus <- sbea(
    method = 'gsea', se = da_output, gs = aer_plus_sig, alpha = 0.05, perm = 0,
    padj.method = 'none'
)
gsRanking(res = enrich_output_plus, signif.only = TRUE)
```

This means anaerobic taxa are enriched in IBD samples? Let's find out:

```{r}
enriched_taxa <- enrich_output_plus$gs$anaerobic
tse_enriched <- da_output[enriched_taxa,]
enriched_taxa_df <- as.data.frame(rowData(tse_enriched))
```


```{r}
enriched_taxa_df$order_names <-
    taxizedb::taxid2name(enriched_taxa_df$order, db = 'ncbi')

enriched_taxa_df |> 
    ggplot(aes(FC, log10(PVAL + 1))) +
    geom_point(aes(color = order_names)) +
    geom_vline(aes(xintercept = 0)) +
    labs(
        title = "Anaerobes are enriched in IBD samples"
    )
```

# Session Information

```{r}
sessionInfo()
```
