---
title: "Enrichment of facultative anaerobes in IBD samples"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugphyzz)
library(curatedMetagenomicData)
library(dplyr)
library(tidySummarizedExperiment)
library(taxPPro)
```

# Expected result

+ Enrichment of anaerobic bacteria in healthy samples (gut).
+ Enrichment of facultative anaerobic bacteria in IBD samples (gut).

# Data

## Abundance data

For this example, let's use the data of the HMP_2019_bidmdb study throug cMD:

```{r, message=FALSE, warning=FALSE}
tse <- curatedMetagenomicData(
    pattern = "HMP_2019_ibdmdb.relative_abundance", 
    dryrun = FALSE, rownames = 'NCBI',
    counts = TRUE
)[[1]]
tse 
```

Use only a subset:

```{r, message=FALSE, warning=FALSE}
rank_cols <- colnames(rowData(tse))
tidy_data <- tidySummarizedExperiment::as_tibble(tse) |> 
    rename(taxon_name = .feature, sample = .sample) 
sample_data <- tidy_data |> 
    select(-rank_cols, -taxon_name, -relative_abundance) |> 
    distinct()
sample_data_split <- split(sample_data, factor(sample_data$disease))
min_n <- min(vapply(sample_data_split, nrow, integer(1)))
samples <- lapply(sample_data_split, function(x) {
    set.seed(1234)
    int_vct <- sample(1:nrow(x), size = min_n)
    x[int_vct,]$sample
}) |> 
    unlist(use.names = FALSE)
tse_subset <- tse[,samples] |> 
    filterTaxa()
tse_subset
```

## Signatures and annotations

```{r}
taxa_data <- tidy_data |> 
    dplyr::filter(taxon_name %in% rownames(tse_subset)) |> 
    select(all_of(c('taxon_name', rank_cols))) |> 
    distinct()
dim(taxa_data)
```

Annotations and microbe signatures come from the bugphyzz package. Since
cMD contains data at the species level, let's use signatures at the species
level. We need to check the completness (number of taxa in cMD samples
annotated in bugphyzz)

```{r, message=FALSE}
aer <- physiologies('aerophilicity')[[1]]
aer_species <- filter(aer, Rank == 'species')
## Check completness
mean(taxa_data$taxon_name %in% aer_species$NCBI_ID) * 100
```

Let's try to increase the completeness with ASR/inheritance propagation using
majority vote:

```{r, message=FALSE, warning=FALSE}
aer_plus <- aer |> 
    propagate()
aer_plus_species <- filter(aer_plus, Rank == 'species')
## Check completeness
mean(taxa_data$taxon_name %in% aer_plus_species$NCBI_ID) * 100
```

Let's get some signatures for aerobic, anaerobic, and facultatively anaerobic:

```{r, message=FALSE, warning=FALSE}
attribute_names <- c('aerobic', 'anaerobic', 'facultatively anaerobic')
aer_sig <- getSignatures(
    aer_plus, tax.id.type = 'NCBI_ID', tax.level = 'species'
)
aer_sig <- aer_sig[names(aer_sig) %in% attribute_names]
lapply(aer_sig, head)
```

For propagated signatures

```{r, message=FALSE, warning=FALSE}
aer_plus_sig <- getSignatures(
    aer_plus, tax.id.type = 'NCBI_ID', tax.level = 'species'
)
aer_plus_sig <- aer_plus_sig[names(aer_plus_sig) %in% attribute_names]
lapply(aer_plus_sig, head)
```


Comparing some numbers

```{r}
df <- data.frame(
    aer = vapply(aer_sig, length, integer(1)),
    aer_plus = vapply(aer_plus_sig, length, integer(1))
)
df
```




I'll use benchadmic.

Get only taxa information for the 'priorInfo' object.


```{r}
taxa_data <- tidy_data |> 
    dplyr::filter(taxon_name %in% rownames(tse_subset)) |> 
    select(all_of(c('taxon_name', rank_cols))) |> 
    distinct()
dim(taxa_data)
```

# Session Information

```{r}
sessionInfo()
```
