---
title: "Analysis of the NYCHANES signatures"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup, message=FALSE}
library(bugphyzzAnalyses)
library(bugsigdbr)
library(bugphyzz)
library(nychanesmicrobiome)
library(purrr)
library(EnrichmentBrowser)
library(tibble)
library(dplyr)
```

Analysis of a previous bsea study: 

# Data

Import data:

```{r, message=FALSE, results='hide'}
## Bugphyzz 
bp <- importBugphyzz(version = 'devel')
aer <- dplyr::filter(bp, Attribute_group == 'aerophilicity')

## BugSigDB
bsdb <- importBugSigDB(version = 'devel')

## TypicalMicrobiomeSignatures
tms <- importTidyTMS()

## NYC HANES TreeSummarizedExperiment
nh <- importNYCHANES()

## NYC HANES - manually curated signatures
biosis_tsv <- system.file('extdata/biosis.tsv', package = 'nychanesmicrobiome')
biosis <- biosis_tsv |> 
  read.table(sep = '\t', header = TRUE) |> 
  dplyr::rename(Genus = X1, aer = X2) |> 
  mutate(aer = tolower(aer)) |> 
  mutate(aer = ifelse(aer == 'f anaerobic', 'facultatively_anaerobic', aer))
```

Create signature sets: 

```{r signature sets}
## bugphyzz sets
bp_taxid_sets <- aer |> 
  getBugphyzzSignatures(tax.id.type = 'NCBI_ID', tax.level = 'genus')

## nychanes sets (biosis file)
nyc_taxname_sets <- map(split(biosis, factor(biosis$aer)), ~ .x$Genus)

## bugphyzz signatures with OTUs
taxids <- rowData(nh)$GenusID
names(taxids) <- rownames(nh)
taxids <- taxids[!is.na(taxids)]
bp_otu_sets <- map(bp_taxid_sets, ~ names(taxids)[which(taxids %in% .x)])
  
## nychanes sets with OTUs  
taxnames <- rowData(nh)$Genus
names(taxnames) <- rownames(nh)
taxnames <- taxnames[!is.na(taxnames)]
nyc_otu_sets <- map(nyc_taxname_sets, ~ names(taxnames)[which(taxnames %in% .x)])

## bugsigdb sets
study_727 <- filter(bsdb, `BSDB ID` %in% c('bsdb:727/1/1', 'bsdb:727/1/2'))
bugsigdb_taxid_sets <- study_727 |> 
  bugsigdbr::getSignatures(tax.id.type = 'ncbi', tax.level = 'mixed') 
names(bugsigdb_taxid_sets) <- sub('^.*_', '', names(bugsigdb_taxid_sets))
bugsigdb_taxname_sets <- study_727 |> 
  bugsigdbr::getSignatures(tax.id.type = 'taxname', tax.level = 'mixed')
names(bugsigdb_taxname_sets) <- sub('^.*_', '', names(bugsigdb_taxname_sets))

## Create backgrounds for bugsigdb sets
oc <- filter(tms, body_site == 'oralcavity', rank == 'genus')
ocbkg_taxid_sets <- list(
  UP = unique(c(bugsigdb_taxid_sets$UP, oc$taxid)),
  DOWN = unique(c(bugsigdb_taxid_sets$DOWN, oc$taxid))
)
ocbkg_taxname_sets <- list(
  UP = unique(c(bugsigdb_taxname_sets$UP, oc$taxon_name)),
  DOWN = unique(c(bugsigdb_taxname_sets$DOWN, oc$taxon_name))
)
```

# Chi-square (bsdb sets)

Contingency tables (bugphyzz)

```{r cont tbl bugphyzz}
aer_subset <- unique(aer[,c('NCBI_ID', 'Attribute')])
df <- data.frame(
    NCBI_ID = c(bugsigdb_taxid_sets$UP, bugsigdb_taxid_sets$DOWN),
    Abundance = c(rep('increased', length(bugsigdb_taxid_sets$UP)), rep('decreased', length(bugsigdb_taxid_sets$DOWN)))
) |> 
  left_join(aer_subset, by = 'NCBI_ID') |> 
  unique()
dups1 <- df$NCBI_ID[which(duplicated(df$NCBI_ID))]
df_1 <- df[!df$NCBI_ID %in% dups1, ]
ct1 <- epitools::epitable(df_1$Attribute, df_1$Abundance)

df_2 <- df |> 
  filter(
    Attribute %in% c('aerophilicity:aerobic', 'aerophilicity:anaerobic')
  )
dups2 <- df_2$NCBI_ID[which(duplicated(df_2$NCBI_ID))]
df_2 <- df_2[!df_2$NCBI_ID %in% dups2, ]
ct2 <- epitools::epitable(df_2$Attribute, df_2$Abundance)
```

Contingency tables (biosis)

```{r cont tbl biosis}
df2 <- data.frame(
    Genus = c(bugsigdb_taxname_sets$UP, bugsigdb_taxname_sets$DOWN),
    Abundance = c(rep('increased', length(bugsigdb_taxname_sets$UP)), rep('decreased', length(bugsigdb_taxname_sets$DOWN)))
) |> 
  left_join(biosis, by = 'Genus') |> 
  unique()

dups3 <- df2$Genus[which(duplicated(df2$Genus))]
df2_1 <- df2[!df2$Genus %in% dups3, ]
ct3 <- epitools::epitable(df2_1$aer, df2_1$Abundance)

df2_2 <- df2 |> 
  filter(
    aer %in% c('aerobic', 'anaerobic')
  )
dups4 <- df2_2$Genus[which(duplicated(df2_2$Genus))]
df2_2 <- df2_2[!df2_2$Genus %in% dups4, ]
ct4 <- epitools::epitable(df2_2$aer, df2_2$Abundance)
```

All attributes (bugphyzz)

```{r}
chi_res1 <- chisq.test(ct1, simulate.p.value = TRUE)
chi_res1
```

Aerobic and anaerobic (bugphyzz)

```{r}
chi_res2 <- chisq.test(ct2, simulate.p.value = TRUE)
chi_res2
```


All attributes (biosis)

```{r}
chi_res3 <- chisq.test(ct3, simulate.p.value = TRUE)
chi_res3
```

aerobic vs anaerobic (biosis)

```{r}
chi_res4 <- chisq.test(ct4, simulate.p.value = TRUE)
chi_res4
```

> Need to check annotations betwwen ct2 and ct4


# Enrihcment of bsdb sets (ORA)


Using bugphyzz signatures:

```{r}
ora_res1 <- runEnrichment(
  bsdb = bsdb,
  bsdb_sets = list(
    `bsdb:727/1/1` = bugsigdb_taxid_sets$DOWN, 
    `bsdb:727/1/2` = bugsigdb_taxid_sets$UP
  ),
  background_sets = list(
    `bsdb:727/1/1` = ocbkg_taxid_sets$DOWN, 
    `bsdb:727/1/2` = ocbkg_taxid_sets$UP
  ),
  bp_sigs = bp_taxid_sets
)
ora_res1
```

Using biosis signatures

Signature 1, Control, aerobic (as expected)
Signature 2, Smokers, anaerobic (as expected)
Sig 1 and 2, fac ana (as expected)

```{r}
ora_res2 <- runEnrichment(
  bsdb = bsdb,
  bsdb_sets = list(
    `bsdb:727/1/1` = bugsigdb_taxname_sets$DOWN, 
    `bsdb:727/1/2` = bugsigdb_taxname_sets$UP
  ),
  background_sets = list(
    `bsdb:727/1/1` = ocbkg_taxname_sets$DOWN, 
    `bsdb:727/1/2` = ocbkg_taxname_sets$UP
  ),
  bp_sigs = nyc_taxname_sets
)
ora_res2
```

# Using NYCHANES original data

## DA

Control 0, smoker 1

```{r}
colData(nh)$GROUP <- ifelse(colData(nh)$smokingstatus == 'Never smoker', 0, 1)
table(nh$GROUP)
```

Control 1, smoker 0

```{r}
nh2 <- nh
colData(nh2)$GROUP <- ifelse(colData(nh2)$smokingstatus == 'Never smoker', 1, 0)
table(nh2$GROUP)
```

DA:

```{r, message=FALSE}
edger <- deAna(
    expr = nh, de.method = 'edgeR', padj.method = 'BH', filter.by.expr = FALSE
)
edger2 <- deAna(
    expr = nh2, de.method = 'edgeR', padj.method = 'BH', filter.by.expr = FALSE
)
```

## ORA

ORA (bugphyzz) control 0, smoker 1:

```{r, message=FALSE}
edger_ora <- sbea(
    method = 'ora', se = edger, gs = bp_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r <- as.data.frame(gsRanking(res = edger_ora, signif.only = FALSE))
r
```
ORA (biosis) control 0, smoker 1:

```{r, message=FALSE}
edger_ora2 <- sbea(
    method = 'ora', se = edger, gs = nyc_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r2 <- as.data.frame(gsRanking(res = edger_ora2, signif.only = FALSE))
r2
```

ORA (bugphyzz) control 1l, smoker 0

```{r, message=FALSE}
edger2_ora <- sbea(
    method = 'ora', se = edger2, gs = bp_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r3 <- as.data.frame(gsRanking(res = edger2_ora, signif.only = FALSE))
r3
```

ORA (biosis) control 1, smoker 0

```{r, message=FALSE}
edger2_ora2 <- sbea(
    method = 'ora', se = edger2, gs = nyc_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r4 <- as.data.frame(gsRanking(res = edger2_ora2, signif.only = FALSE))
r4
```

## GSEA

GSEA (bugphyzz) control 0, smoker 1:

```{r, message=FALSE}
edger_gsea <- sbea(
    method = 'gsea', se = edger, gs = bp_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r5 <- as.data.frame(gsRanking(res = edger_gsea, signif.only = FALSE))
r5
```

GSEA (biosis) control 0, smoker 1:

```{r, message=FALSE}
edger_gsea2 <- sbea(
    method = 'gsea', se = edger, gs = nyc_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r6 <- as.data.frame(gsRanking(res = edger_gsea2, signif.only = FALSE))
r6
```

GSEA (bugphyzz) control 1l, smoker 0

```{r, message=FALSE}
edger2_gsea <- sbea(
    method = 'gsea', se = edger2, gs = bp_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r8 <- as.data.frame(gsRanking(res = edger2_gsea, signif.only = FALSE))
r8
```

GSEA (biosis) control 1, smoker 0

```{r, message=FALSE}
edger2_gsea2 <- sbea(
    method = 'gsea', se = edger2, gs = nyc_otu_sets, alpha = 0.1, perm = 0,
    padj.method = 'BH'
)
r8 <- as.data.frame(gsRanking(res = edger2_gsea2, signif.only = FALSE))
r8
```

# Session information

```{r}
sessioninfo::session_info()
```
