---
title: "Some numbers"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message=FALSE}
library(taxPPro)
library(bugphyzz)
library(dplyr)
library(ggplot2)
```

## Comparison of a categorical and a continuous attribute

Attribute with categorical values (aerophilicity):

```{r, message=FALSE, warning=FALSE}
aer <- physiologies('aerophilicity')[[1]]
aer_asr <- propagate(aer, max.tax.level = 'family', direction = 'upstream')
aer_inh <- propagate(aer, max.tax.level = 'family',  direction = 'downstream')
aer_prop <- propagate(aer, max.tax.level = 'family', direction = 'both')
aer_data <- list(
    original = aer, asr = aer_asr, inheritance = aer_inh,
    `asr + inheritance` = aer_prop
)
aer_data <- lapply(aer_data, function(x) {
    x$NCBI_ID <- as.character(x$NCBI_ID)
    x$Parent_NCBI_ID <- as.character(x$Parent_NCBI_ID)
    x
})
aer_data <- aer_data |> 
    bind_rows(.id = 'method')
aer_data_summary <- aer_data |> 
  count(method, Evidence, Rank) |> 
  filter(Rank %in% c('genus', 'species', 'strain')) |> 
  mutate(
    method = factor(
      method, c('original', 'asr', 'inheritance', 'asr + inheritance')
    )
  ) 
```

Plot numbers

```{r}
aer_data_summary |> 
  ggplot(aes(method, n)) +
  geom_col(aes(fill = Evidence)) + 
  labs(
    title = 'Comparison of propagation methods for aerophilicity',
    subtitle = 'Categorical; max.tax.level = genus'
  ) +
  facet_wrap(~Rank, scales = 'free') +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) 
```


In the case of the areophilicity dataset, the number of genera increases
from about 1200 to 2000, when using asr. The number of species increases from
about 10,000 to almost 15,000. There is no much change in the number of strains.
The number strains goes from 35,000 to about 40,000.

The low number of species obtained through ASR is due to the low levels of 
strains present in the original data (right panel). However, it seems that
these few species make a considerably difference when 

Attribute with continuous values (optimal ph):

```{r, message=FALSE, warning=FALSE}
ph <- physiologies('optimal ph')[[1]]
ph_asr <- propagate(ph, max.tax.level = 'genus', direction = 'upstream')
ph_inh <- propagate(ph, max.tax.level = 'genus', direction = 'downstream')
ph_prop <- propagate(ph, max.tax.level = 'genus', direction = 'both')
ph_data <- list(
    original = ph, asr = ph_asr, inheritance = ph_inh,
    `asr + inheritance` = ph_prop
)
ph_data <- lapply(ph_data, function(x) {
    x$NCBI_ID <- as.character(x$NCBI_ID)
    x$Parent_NCBI_ID <- as.character(x$Parent_NCBI_ID)
    x
})
ph_data <- ph_data |> 
    bind_rows(.id = 'method')
ph_data_summary <- ph_data |> 
  count(method, Evidence, Rank) |> 
  filter(Rank %in% c('genus', 'species', 'strain')) |> 
  mutate(
    method = factor(
      method, c('original', 'asr', 'inheritance', 'asr + inheritance')
    )
  ) 
```

Plot numbers:

```{r}
ph_data_summary |> 
  ggplot(aes(method, n)) +
  geom_col(aes(fill = Evidence)) + 
  labs(
    title = 'Comparison of propagation methods for aerophilicity',
    subtitle = 'Categorical; max.tax.level = genus'
  ) +
  facet_wrap(~ Rank, scales = 'free') +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) 

```

This dataset didn't have any genera. So, all genera come from ASR. The increase
in the number of taxa at the species and strain levels seems considerable.

## Session information

```{r}
sessionInfo()
```
