---
title: "Odds Ratio - Up vs Down"
output:
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bugphyzzAnalyses)
library(bugphyzz)
library(bugsigdbr)
library(dplyr)
library(purrr)
library(tidyr)
library(ComplexHeatmap)
```

## Import BugPhyzz

```{r, warning=FALSE, message=FALSE}
bp <- importBugphyzz()
bp_sigs_gn <- map(bp, ~ {
    makeSignatures(
        dat = .x, tax_id_type = "NCBI_ID", tax_level = "genus", min_size = 5
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null) |> 
    map(as.character)

bp_sigs_sp <- map(bp, ~ {
    makeSignatures(
        dat = .x, tax_id_type = "NCBI_ID", tax_level = "species", min_size = 5
    )
}) |> 
    list_flatten(name_spec = "{inner}") |> 
    discard(is.null) |> 
    map(as.character)
```

## Import data

Use "case-control" study design only:

```{r, message=FALSE}
## 10.5281/zenodo.10407666 v1.2.0
## "10.5281/zenodo.6468009" v1.1.0
bsdb <- importBugSigDB(version = "10.5281/zenodo.10407666" , cache = FALSE)
# bsdb <- importBugSigDB(version = "devel" , cache = FALSE)
bsdb <- bsdb |> 
    filter(`Host species` == "Homo sapiens") |> 
    filter(!is.na(`Abundance in Group 1`)) |> 
    filter(!is.na(`Body site`)) |> 
    filter(`Study design` == "case-control" ) |> 
    mutate(
        `BSDB ID` = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment), "/",
            sub("Signature ", "", `Signature page name`)
        )
    ) |> 
    mutate(
        custom_id = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment)
        )
    ) |> 
    relocate(custom_id, `BSDB ID`)
dim(bsdb)
```

Create a list with signature ids (BSDB ID) that have both UP and DOWN 
outcomes per study:

```{r}
## Columns that make BSDB ID: Study, Experiment, Signature page name
l <- bsdb |> 
    select(
        Study, Experiment, `Signature page name`, `Abundance in Group 1`
    ) |> 
    mutate(
        xcol = paste0(`Abundance in Group 1`, " (", `Signature page name`, ")")
    ) |> 
    group_by(Study, Experiment) |> 
    mutate(ycol = sort(paste0(xcol, collapse = "; "))) |> 
    mutate(n_sig = n()) |> 
    ungroup() |> 
    select(-xcol) |> 
    filter(n_sig == 2) |> 
    filter(grepl("decreased", ycol), grepl("increased", ycol)) |> 
    filter(grepl("Signature 1", ycol), grepl("Signature 2", ycol)) |> 
    mutate(
        `BSDB ID` = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment), "/",
            sub("Signature ", "", `Signature page name`)
        )
    ) |> 
    mutate(
        custom_id = paste0(
            "bsdb:",
            sub("Study ", "", Study), "/",
            sub("Experiment ", "", Experiment)
        )
    ) |> 
    relocate(custom_id, `BSDB ID`) |> 
    {\(y) split(y, y$`Abundance in Group 1`)}() |> 
    map(~ pull(arrange(.x, custom_id), `BSDB ID`))
map(l, length)
```

## Genus

Create signatures for both decrease and increase at the **genus** level. 
Only keep the signature names (stored in a data.frame) that have enough taxa in
both increased and decreased sets.

```{r}
lg <- map(l, ~ {
    bsdb |> 
        filter(`BSDB ID` %in% .x) |> 
        getSignatures(
            tax.id.type = "ncbi", tax.level = "genus", exact.tax.level = FALSE,
            min.size = 5
        ) |> 
        keep(~ length(.x) >= 5)
})
df_gn_1 <- data.frame(
    x = paste0(
        sub("/\\d_.*", "", names(lg$decreased)), "---", names(lg$decreased)
    )
) |> 
    separate(col = "x", into = c("custom_id", "sig_down"), sep = "---")

df_gn_2 <- data.frame(
    x = paste0(
        sub("/\\d_.*", "", names(lg$increased)), "---", names(lg$increased)
    )
) |> 
    separate(col = "x", into = c("custom_id", "sig_up"), sep = "---")

df_gn <- full_join(df_gn_1, df_gn_2, by = "custom_id") |> 
    drop_na() |> 
    mutate(
        check = sub("bsdb:\\d+/\\d+/\\d+_(.*)_UP", "\\1", sig_up) ==
           sub("bsdb:\\d+/\\d+/\\d+_(.*)_DOWN", "\\1", sig_down)
    )
dim(df_gn)    
```

## Species

Create signatures for both decrease and increase at the **species** level.
Only keep the signature names (stored in a data.frame) that have enough taxa in
both increased and decreased sets.

```{r}
ls <- map(l, ~ {
    bsdb |> 
        filter(`BSDB ID` %in% .x) |> 
        getSignatures(
            tax.id.type = "ncbi", tax.level = "species",
            exact.tax.level = FALSE, min.size = 5
        ) |> 
        keep(~ length(.x) >= 5)
})
df_sp_1 <- data.frame(
    x = paste0(
        sub("/\\d_.*", "", names(ls$decreased)), "---", names(ls$decreased)
    )
) |> 
    separate(col = "x", into = c("custom_id", "sig_down"), sep = "---")

df_sp_2 <- data.frame(
    x = paste0(
        sub("/\\d_.*", "", names(ls$increased)), "---", names(ls$increased)
    )
) |> 
    separate(col = "x", into = c("custom_id", "sig_up"), sep = "---")

df_sp <- full_join(df_sp_1, df_sp_2, by = "custom_id") |> 
    drop_na() |> 
    mutate(
        check = sub("bsdb:\\d+/\\d+/\\d+_(.*)_UP", "\\1", sig_up) ==
            sub("bsdb:\\d+/\\d+/\\d+_(.*)_DOWN", "\\1", sig_down)
    )
dim(df_sp)
```

## Get lists of BSDB signatures

```{r}
sigs_list_gn <- list()
sigs_list_gn[["up"]] <- lg$increased[df_gn$sig_up]
sigs_list_gn[["down"]] <- lg$decreased[df_gn$sig_down]

sigs_list_sp <- list()
sigs_list_sp[["up"]] <- ls$increased[df_sp$sig_up]
sigs_list_sp[["down"]] <- ls$decreased[df_sp$sig_down]
```

## Odds ratio - function

```{r}
or <- function(x, y, s) {
    a <- sum(x %in% s)
    b <- sum(!x %in% s)
    c <- sum(y %in% s)
    d <- sum(!y %in% s)
    ct <- matrix(
        data = c(a, b, c, d), byrow = TRUE, nrow = 2, ncol = 2,
        dimnames = list(exposure = c("yes", "no"), event = c("yes", "no"))
    )
    # suppressWarnings({
    #     res <- epitools::oddsratio.wald(ct)
    # })
    p_value = fisher.test(x = ct, alternative = "two.sided")$p.value
    
    if (any(ct == 0)) {
        HA <- "*"
        suppressWarnings({
            # res <- epitools::oddsratio.small(ct)
            # p_value = fisher.test(x = ct + 0.5, alternative = "two.sided")$p.value
            res <- epitools::oddsratio.wald(ct + 0.5) # I get a lot of caution messages about not using this funcion
        })
    } else {
        HA <- ""
        suppressWarnings({
            # p_value = fisher.test(x = ct, alternative = "two.sided")$p.value
            res <- epitools::oddsratio.wald(ct)
        })
    }
    
    output <- data.frame(
       odds_ratio = round(res$measure[2, "estimate"], 2),
       lower_ci = round(res$measure[2, "lower"], 2),
       upper_ci = round(res$measure[2, "upper"], 2),
       # p_value = res$p.value[2, "fisher.exact"],
       p_value = p_value,
       # method = method,
       HA = HA,
       up_annotated = a,
       up_noAnnotated = b,
       down_annotated = c,
       down_noAnnotated = d
    )
    return(output)
}
```

## Calculate odds ratio and p values

```{r}
fname <- system.file(
    'extdata', 'condition2category.tsv', package = 'bugphyzzAnalyses',
    mustWork = TRUE
)
cond2cat <- read.table(fname, header = TRUE, sep = '\t')
```

```{r}
gn <- vector("list", length(sigs_list_gn$up) * length(bp_sigs_gn))
counter <- 1
for (i in seq_along(sigs_list_gn$up)) {
    for(j in seq_along(bp_sigs_gn)) {
        odds_ratio <- or(
            x = sigs_list_gn$up[[i]],
            y = sigs_list_gn$down[[i]],
            s = bp_sigs_gn[[j]]
        )
        names(gn)[counter] <- paste0(
            names(sigs_list_gn$up)[i], "---", names(bp_sigs_gn)[j], "---genus"
        )
        gn[[counter]] <- odds_ratio
        counter <- counter + 1
    }
}
gn_df <- gn |> 
    bind_rows(.id = "name") |> 
    separate(
        col = "name", into = c("bsdb_sig", "bp_sig", "rank"), sep = "---"
    ) |> 
    mutate(
        `BSDB ID` = sub("^(bsdb:\\d+/\\d+/\\d+)_.*$", "\\1", bsdb_sig),
        bp_sig = sub("^bugphyzz:(.*)$", "\\1", bp_sig)
    ) |> 
    left_join(select(bsdb, `BSDB ID`, Condition), by = "BSDB ID") |> 
    relocate(`BSDB ID`, Condition) |> 
    relocate(bsdb_sig, .after = down_noAnnotated) |> 
    arrange(-odds_ratio) |> 
    mutate(Condition = tolower(Condition)) |> 
    left_join(cond2cat, by = "Condition") |> 
    relocate(Category, .after = `BSDB ID`) |> 
    filter(!is.na(Category))
dim(gn_df)
```

```{r}
sp <- vector("list", length(sigs_list_sp$up) * length(bp_sigs_sp))
counter <- 1
for (i in seq_along(sigs_list_sp$up)) {
    for(j in seq_along(bp_sigs_sp)) {
        odds_ratio <- or(
            x = sigs_list_sp$up[[i]],
            y = sigs_list_sp$down[[i]],
            s = bp_sigs_sp[[j]]
        )
        names(sp)[counter] <- paste0(
            names(sigs_list_sp$up)[i], "---", names(bp_sigs_sp)[j], "---species"
        )
        sp[[counter]] <- odds_ratio
        counter <- counter + 1
    }
}
sp_df <- sp |> 
    bind_rows(.id = "name") |> 
    separate(
        col = "name", into = c("bsdb_sig", "bp_sig", "rank"), sep = "---"
    ) |> 
    mutate(
        `BSDB ID` = sub("^(bsdb:\\d+/\\d+/\\d+)_.*$", "\\1", bsdb_sig),
        bp_sig = sub("^bugphyzz:(.*)$", "\\1", bp_sig)
    ) |> 
    left_join(select(bsdb, `BSDB ID`, Condition), by = "BSDB ID") |> 
    relocate(`BSDB ID`, Condition) |> 
    relocate(bsdb_sig, .after = down_noAnnotated) |> 
    arrange(-odds_ratio) |> 
    mutate(Condition = tolower(Condition)) |> 
    left_join(cond2cat, by = "Condition") |> 
    relocate(Category, .after = `BSDB ID`) |> 
    filter(!is.na(Category))
dim(sp_df)
```

## Merge gn and sp

```{r}
mainDF <- bind_rows(gn_df, sp_df)
dim(mainDF)
```

## odds ratio - genus - table up

```{r}
mainDF |> 
    filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    arrange(-odds_ratio) |> 
    {\(y) myDataTable(y, 30)}()
```

## Odds ratio - genus - heatmap up

```{r, fig.height=9}
gn_mat_up <- mainDF |> 
    filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

color_fun <- function(x) {
    circlize::colorRamp2(
        breaks = c(0, seq(1, max(x))),
        colors = c("white", viridis::viridis(max(x)))
    )
}

gn_hp_up <- Heatmap(
    matrix = gn_mat_up,
    col = color_fun(gn_mat_up), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(gn_mat_up),
        gp = gpar(fontsize = 12)
    ),
)
gn_hp_up
```

## odds ratio - genus - table down

```{r}
gn_df |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    arrange(-odds_ratio) |> 
    {\(y) myDataTable(y, length(y))}()
```

## odds ratio - genus - heatmap down

```{r fig.height=9}
gn_mat_down <- mainDF |> 
    filter(rank == "genus") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()
gn_hp_down <- Heatmap(
    matrix = gn_mat_down,
    col = color_fun(gn_mat_down), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(gn_mat_down),
        gp = gpar(fontsize = 12)
    ),
)
gn_hp_down
```

## odds ratio - species - table up

```{r}
mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    arrange(-odds_ratio) |> 
    {\(y) myDataTable(y, nrow(y))}()
```

## Odds ratio - species - heatmap up

```{r, fig.height=9}
sp_mat_up <- mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio > 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

sp_hp_up <- Heatmap(
    matrix = sp_mat_up,
    col = color_fun(sp_mat_up), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(sp_mat_up),
        gp = gpar(fontsize = 12)
    ),
)
sp_hp_up
```

## Odds ratio - species - table down

```{r}
mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    arrange(-odds_ratio) |> 
    {\(y) myDataTable(y, length(y))}()
```

## Odds ratio - species - heatmap down

```{r, fig.height=9}
sp_mat_down <- mainDF |> 
    filter(rank == "species") |> 
    filter(p_value < 0.05, odds_ratio < 1) |> 
    count(Category, bp_sig) |>
    pivot_wider(
        names_from = "Category", values_from = "n", values_fill = 0
    ) |>
    tibble::column_to_rownames(var = "bp_sig") |>
    as.matrix()

sp_hp_down <- Heatmap(
    matrix = sp_mat_down,
    col = color_fun(sp_mat_down), 
    border = TRUE,
    show_row_dend = FALSE, show_column_dend = FALSE,
    row_names_side = "left",
    name = "# sigs",
    rect_gp = gpar(col = "gray50", lwd = 1), 
        row_names_max_width = max_text_width(
        rownames(sp_mat_down),
        gp = gpar(fontsize = 12)
    ),
)
sp_hp_down
```

## Session information

```{r}
sessioninfo::session_info()
```
