---
title: "Results of 10-fold cross-validation of predicted annotations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, message=FALSE, echo=FALSE}
library(DT)
library(dplyr)
```

## Import validation data

Validation scripts and outputs are in: https://github.com/waldronlab/taxPProValidation/tree/main

### Columns description

| Column | Description |
| ------ | ----------- |
| **method** | The R package used for the analysis. Either phytools or castor. ltp stands for Living Tree Project. Phytools was used for discrete attributes. castor was used for numeric attributes. |
| **rank** | Rank of the input annotated taxa used. 'all' means no filter, so the data could contain genus, species, and strains, but possibly just one of two of these ranks.|
| **physiology** | Name of the physiology in bugphyzz. |
| **attribute** | Name of the attribute. In the case of numeric attributes, it's the same name of the physiology. |
| **mcc_mean and mcc_sd ** | Mean and standard deviation of the Mathews correlation coefficient values. It's the result of 10-fold cross-validation. These values only apply to discrete attributes. |
| **r_squared_mean and r_squated_sd** | Mean and standard deviation of the R-square values. It's the result of 10-fold cross-validation. These values only apply to numeric attributes.|
| **ltp_and_bp** | Number of annotations in both the LTP tree and a given subset of a bughyzz dataset (per rank and attribute). |
| **ltp_and_bp_phys** | Number of annotations in both the LTP tree and a given set of a bugphyzz physiology (adding up all attributes).|
| **bugphyzz** | Number of taxa with annotations in a given subset of a bugphyzz dataset (per rank and attribute). |
| **ltp**  | total number of tips in the LTP tree. These tips include a mix of genera, species, and strains.|

```{r import validation data, echo=FALSE}
url <- "https://raw.githubusercontent.com/waldronlab/taxPProValidation/main/validation_summary.tsv"
dat <- read.table(file = url, header = TRUE, sep = "\t", row.names = NULL) |> 
    arrange(method, rank, physiology, attribute) |> 
    group_by(method, rank, physiology) |> 
    mutate(ltp_and_bp_phys = sum(ltp_and_bugphyzz)) |> 
    ungroup() |> 
    rename(ltp_and_bp = ltp_and_bugphyzz) |> 
    relocate(ltp_and_bp_phys, .after = ltp_and_bp)
        
datatable(
    data = dat, filter = "top", rownames = FALSE,
    options = list(
        pageLength = 100,
        searchCols = list(
            NULL, # first column
            list(search = 'all'), # secound column
            NULL # other columns
        )

    )
)
```

## Session information

```{r, echo=FALSE}
sessioninfo::session_info()
```


