---
title: "Results of 10-fold cross-validation of predicted annotations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, echo=FALSE}
library(DT)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(bugphyzzAnalyses)
```

# Import validation data

Validation scripts and outputs are in: https://github.com/waldronlab/taxPProValidation/tree/main

### Columns description

| Column | Description |
| ------ | ----------- |
| **method** | The R package used for the analysis. Either phytools or castor. ltp stands for Living Tree Project. Phytools was used for discrete attributes. castor was used for numeric attributes. |
| **rank** | Rank of the input annotated taxa used. 'all' means no filter, so the data could contain genus, species, and strains, but possibly just one of two of these ranks.|
| **physiology** | Name of the physiology in bugphyzz. |
| **attribute** | Name of the attribute. In the case of numeric attributes, it's the same name of the physiology. |
| **mcc_mean and mcc_sd ** | Mean and standard deviation of the Mathews correlation coefficient values. It's the result of 10-fold cross-validation. These values only apply to discrete attributes. |
| **r_squared_mean and r_squated_sd** | Mean and standard deviation of the R-square values. It's the result of 10-fold cross-validation. These values only apply to numeric attributes.|
| **ltp_and_bp** | Number of annotations in both the LTP tree and a given subset of a bughyzz dataset (per rank and attribute). |
| **ltp_and_bp_phys** | Number of annotations in both the LTP tree and a given set of a bugphyzz physiology (adding up all attributes).|
| **bugphyzz** | Number of taxa with annotations in a given subset of a bugphyzz dataset (per rank and attribute). |
| **ltp**  | total number of tips in the LTP tree. These tips include a mix of genera, species, and strains.|
| **nsti_mean and nsti_sd**  | Mean and standard deviation of NSTI values of unknwon tips. Lower is better. |

```{r import validation data, echo=FALSE}
url <- "https://raw.githubusercontent.com/waldronlab/taxPProValidation/main/validation_summary.tsv"
dat <- read.table(file = url, header = TRUE, sep = "\t", row.names = NULL) |> 
    arrange(method, rank, physiology, attribute)
    
# dat |>
#     DT::datatable(
#         rownames = FALSE,
#         extensions = "Buttons",
#         filter = "top",
#         options = list(
#             dom = 'lBfrtip',
#             paging = TRUE,
#             pageLength = nrow(dat),
#             searchCols = list(
#                 NULL, # first column
#                 list(search = 'all'), # secound column
#                 NULL # other columns
#             ),
#             buttons = list(
#                 list(extend = "copy", title = NULL),
#                 list(extend = "print", title = NULL),
#                 list(extend = "csv", title = NULL)
#             )
#         )
# )
```


```{r}
output <- dat |> 
    filter(rank == "all") |> 
    select(
        method,
        Attribute = physiology,
        `Attribute value` = attribute,
        `MCC mean` = mcc_mean,
        `MCC sd` = mcc_sd,
        `R-squared mean` = r2_mean,
        `R-squared sd` = r2_sd,
        `No. annotated tips per attribute value` = ltp_bp,
        `No. annotated tips per attribute` = ltp_bp_phys,
        `Percent of annotated tips per attribute value` = ltp_bp_per,
        `Percent of annotated tips per attribute` = ltp_bp_phys_per
    ) |> 
    arrange(
        -`MCC mean`, -`R-squared mean`, method, Attribute, `Attribute value`) |> 
    select(-method)
myDataTable(output, page_len = nrow(output))
```


# Ontology data

```{r}
fname <- system.file(
    "extdata", "attributes.tsv", package = "bugphyzz"
)
o <- readr::read_tsv(fname, show_col_types = FALSE)
```


```{r}
o2 <- o |> 
    mutate(ontology = gsub("(?<!CO)_(?!(\\d+:\\d+))", ":", ontology, perl = TRUE)) |>
    mutate(ontology = sub("^(.*):.*$", "\\1", ontology)) |> 
    select(ontology) |> 
    filter(!is.na(ontology)) |> 
    arrange(ontology) |> 
    distinct()
o3 <- o2 |> 
    mutate(
        Description = case_when(
            ontology == "APO" ~ "obophenotype/ascomycete-phenotype-ontology",
            ontology == "ARO" ~ "Antibiotic Resistance Ontology",
            ontology == "BTO" ~ "The BRENDA Tissue Ontology",
            ontology == "CO_320" ~ "The Planteome Project",
            ontology == "CO_331" ~ "The Planteome Project",
            ontology == "CO_345" ~ "The Planteome Project",
            ontology == "CO_357" ~ "The Planteome Project",
            ontology == "ECOCORE" ~ "An ontology of core ecological entities",
            ontology == "EFO" ~ "Experimental Factor Ontology",
            ontology == "EHDAA2" ~ "Ontobee/Human developmental anatomy, abstract",
            ontology == "ENM" ~ "eNanoMapper ontology",
            ontology == "ENVO" ~ "The Environment Ontology",
            ontology == "ERO" ~ "Eagle-I Research Resource Ontology",
            ontology == "FMA" ~ "The Foundational Model of Anatomy",
            ontology == "FOODON" ~ "Ontobee/Food Ontology",
            ontology == "GO" ~ "Gene Ontology",
            ontology == "IDO" ~ "Infectious Disease Ontology",
            ontology == "IDOMAL" ~ "Malaria Ontology",
            ontology == "MCO" ~ "Microbial Conditions Ontology",
            ontology == "MICRO" ~ "Ontology of Prokaryotic Phenotypic and Metabolic Characters",
            ontology == "MONDO" ~ "Mondo Disease Ontology",
            ontology == "MP" ~ "The Mammalian Phenotype Ontology",
            ontology == "NCBITaxon" ~ "NCBI organismal classification",
            ontology == "NCIT" ~ "NCI Thesaurus OBO Edition",
            ontology == "OBI" ~ "Ontology for Biomedical Investigations",
            ontology == "OHMI" ~ "OHMI: Ontology of Host-Microbiome Interactions",
            ontology == "OMIT" ~ "Ontology for MIRNA Target",
            ontology == "OMP" ~ "Ontology of Microbial Phenotypes",
            ontology == "Orphanet" ~ "Orphanet",
            ontology == "PATO" ~ "PATO - the Phenotype And Trait Ontology",
            ontology == "PHIPO" ~ "Pathogen Host Interactions Phenotype Ontology",
            ontology == "PO" ~ "Plant Ontology",
            ontology == "SIO" ~ "Ontobee/Semanticscience Integrated Ontology",
            ontology == "SNOMED" ~ "SNOMED CT (International Edition)",
            ontology == "SPD" ~ "The Spider Anatomy Ontology",
            ontology == "SYMP" ~ "Symptom Ontology",
            ontology == "UBERON" ~ "Uber-anatomy ontology",
            ontology == "UPa" ~ "Unipathway",
            ontology == "XCO" ~ "Experimental condition ontology",
            ontology == "ZFA" ~ "Zebrafish Anatomy Ontology (ZFA)"
        )
    ) |> 
    mutate(
        ontology = case_when(
            grepl("^(CO_).*$", ontology) ~ "CO",
            TRUE ~ ontology
    )) |> 
    distinct() |>
    dplyr::rename(Ontology = ontology)
myDataTable(o3)
```

# Session information

```{r, echo=FALSE}
sessioninfo::session_info()
```
