---
title: "Results of 10-fold cross-validation of predicted annotations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, message=FALSE, echo=FALSE}
library(DT)
library(dplyr)
library(ggplot2)
library(ggrepel)
```

## Import validation data

Validation scripts and outputs are in: https://github.com/waldronlab/taxPProValidation/tree/main

### Columns description

| Column | Description |
| ------ | ----------- |
| **method** | The R package used for the analysis. Either phytools or castor. ltp stands for Living Tree Project. Phytools was used for discrete attributes. castor was used for numeric attributes. |
| **rank** | Rank of the input annotated taxa used. 'all' means no filter, so the data could contain genus, species, and strains, but possibly just one of two of these ranks.|
| **physiology** | Name of the physiology in bugphyzz. |
| **attribute** | Name of the attribute. In the case of numeric attributes, it's the same name of the physiology. |
| **mcc_mean and mcc_sd ** | Mean and standard deviation of the Mathews correlation coefficient values. It's the result of 10-fold cross-validation. These values only apply to discrete attributes. |
| **r_squared_mean and r_squated_sd** | Mean and standard deviation of the R-square values. It's the result of 10-fold cross-validation. These values only apply to numeric attributes.|
| **ltp_and_bp** | Number of annotations in both the LTP tree and a given subset of a bughyzz dataset (per rank and attribute). |
| **ltp_and_bp_phys** | Number of annotations in both the LTP tree and a given set of a bugphyzz physiology (adding up all attributes).|
| **bugphyzz** | Number of taxa with annotations in a given subset of a bugphyzz dataset (per rank and attribute). |
| **ltp**  | total number of tips in the LTP tree. These tips include a mix of genera, species, and strains.|
| **nsti_mean and nsti_sd**  | Mean and standard deviation of NSTI values of unknwon tips. Lower is better. |

```{r import validation data, echo=FALSE}
url <- "https://raw.githubusercontent.com/waldronlab/taxPProValidation/main/validation_summary.tsv"
dat <- read.table(file = url, header = TRUE, sep = "\t", row.names = NULL) |> 
    arrange(method, rank, physiology, attribute) |>
    group_by(method, rank, physiology)
        
 datatable(
     data = dat, filter = "top", rownames = FALSE,
     options = list(
         pageLength = 100,
         searchCols = list(
             NULL, # first column
             list(search = 'all'), # secound column
             NULL # other columns
         )
 
     )
 )
```


## Some plots

```{r, echo=FALSE}
dat <- dat |> 
    mutate(
        attribute_type = case_when(
            method == "castor-ltp" ~ "numeric",
            method == "phytools-ltp" & physiology == attribute ~ "binary",
            method == "phytools-ltp" & physiology != attribute ~ "multistate",
            TRUE ~ NA
        )
    )
```


### Binary

```{r}
p_binary <- dat |> 
    filter(attribute_type == "binary") |> 
    ggplot(aes(ltp_bp_phys, nsti_mean)) +
    geom_point(aes(color = mcc_mean), size = 3) +
    geom_text_repel(aes(label = paste0(physiology, " (", mcc_mean, ")"))) +
    facet_wrap(~rank) +
    scale_color_viridis_c(option = "C") +
    theme_bw()
p_binary
```

### Numeric

```{r}
p_numeric <- dat |> 
    mutate(r2_mean = ifelse(r2_mean < 0, 0, r2_mean)) |> 
    filter(attribute_type == "numeric") |> 
    ggplot(aes(ltp_bp_phys, nsti_mean)) +
    geom_point(aes(color = r2_mean), size = 3) +
    geom_text_repel(aes(label = paste0(physiology, " (", r2_mean, ")"))) +
    facet_wrap(~rank) +
    scale_color_viridis_c(option = "C") +
    theme_bw()
p_numeric
```

### Multistate

```{r}
p_multi <- dat |> 
    filter(attribute_type == "multistate") |> 
    group_by(physiology) |> 
    mutate(
        mcc_mean = round(mean(mcc_mean), 2),
        nsti_mean = round(mean(nsti_mean), 2)
    ) |>
    ungroup() |> 
    select(physiology, mcc_mean, nsti_mean, ltp_bp_phys, rank) |> 
    distinct() |> 
    ggplot(aes(ltp_bp_phys, nsti_mean)) +
    geom_point(aes(color = mcc_mean), size = 3) +
    geom_text_repel(aes(label = paste0(physiology, " (", mcc_mean, ")"))) +
    facet_wrap(~rank) +
    scale_color_viridis_c(option = "C") +
    theme_bw()
p_multi
```



## Session information

```{r, echo=FALSE}
sessioninfo::session_info()
```


